---
title: "Hands-on Exercise 4D"
author: "Alvin Ong Wei Ming"
date: "May 09, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: true #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

# Funnel plot for visualisation

## How can funnel plot be used.

Funnel plot is a specially designed data visualisation for conducting unbiased comparison between outlets, stores or business entities.

## packages used

-   [readr](https://cran.r-project.org/web/packages/readr/index.html) for importing csv into R.
-   [FunnelPlotR](https://cran.r-project.org/web/packages/FunnelPlotR/index.html) for creating funnel plot.
-   [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html) for creating funnel plot manually.
-   [knitr](https://cran.r-project.org/web/packages/knitr/index.html) for building static html table.
-   [plotly](https://cran.r-project.org/web/packages/plotly/index.html) for creating interactive funnel plot.

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

COVID-19_DKI_Jakarta will be used. The data was downloaded from Open Data Covid-19 Provinsi DKI Jakarta portal. For this hands-on exercise, we are going to compare the cumulative COVID-19 cases and death by sub-district (i.e. kelurahan) as at 31st July 2021, DKI Jakarta.

```{r}
covid19 <- read_csv("COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)

head(covid19)
```

## FunnelPlotR methods: Basic Plot Example

[FunnelPlotR](https://nhs-r-community.github.io/FunnelPlotR/) package uses ggplot to generate funnel plots. It requires a numerator (events of interest), denominator (population to be considered) and group. The key arguments selected for customisation are:

-   limit: plot limits (95 or 99).
-   label_outliers: to label outliers (true or false).
-   Poisson_limits: to add Poisson limits to the plot.
-   OD_adjust: to add overdispersed limits to the plot.
-   xrange and yrange: to specify the range to display for - axes, acts like a zoom function.
-   Other aesthetic components such as graph title, axis labels etc.

::: callout-note
A funnel plot expects the denominator to be the "opportunity" for an event to happen (the larger population at risk) and the numerator to be the "event" itself (the smaller count)
:::

:::: panel-tabset
## Basic Plot

::: callout-note
-   group in this function is different from the scatterplot. Here, it defines the level of the points to be plotted i.e. Sub-district, District or City. If Cityc is chosen, there are only six data points.

-   By default, data_typeargument is “SR”.

-   limit: Plot limits, accepted values are: 95 or 99, corresponding to 95% or 99.8% quantiles of the distribution.
:::

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  title = "Funnel Plot of COVID-19 Positives vs Deaths"
)
```

## Makeover 1

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)
```

## Makeover 2

Things to learn from the code chunk above.

-   label = NA argument is to removed the default label outliers feature.
-   title argument is used to add plot title.
-   x_label and y_label arguments are used to add/edit x-axis and y-axis titles.

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```
::::

## Computing the basic derived fields with ggplot2

To plot the funnel plot from scratch, we need to derive cumulative death rate and standard error of cumulative death rate.

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)

```

::: callout-note
-   rate (variable): Observed rate

-   rate.se (variable): This uses the binomial distribution formula to calculate the Standard Error of the rate

-   finally we filter out 0 for any sub-districts that reported zero deaths.
:::

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

::: callout-note
The reason you use weighted.mean instead of a simple mean() is to create a "fair" baseline that recognizes not all data points are equally reliable.

1/df\$rate.se\^2 refers to inverse-variance weighting

It gives a much higher weight to sub-districts with high precision (small SE) and less weight to sub-districts with low precision (large SE). Hence we do not want the arithmetic mean but using our defined mean function
:::

## if we do not calculate a weighted mean

```{r}
#| echo: false

simple.mean <- mean(df$rate)

# 2. Generate smooth funnel lines based on the weighted mean
# We create a sequence of 'Positive' population sizes to draw the curve
funnel_bounds <- data.frame(Positive = seq(min(df$Positive), max(df$Positive), length.out = 100)) %>%
  mutate(

    expected_se = sqrt((fit.mean * (1 - fit.mean)) / Positive),

    limit_95_upper = fit.mean + 1.96 * expected_se,
    limit_95_lower = fit.mean - 1.96 * expected_se,

    limit_99_upper = fit.mean + 3.09 * expected_se,
    limit_99_lower = fit.mean - 3.09 * expected_se
  )

# 3. Plot the Funnel Chart
ggplot() +
  # Draw the Funnel boundary lines
  geom_line(data = funnel_bounds, aes(x = Positive, y = limit_95_upper), linetype = "dashed", color = "gray50") +
  geom_line(data = funnel_bounds, aes(x = Positive, y = limit_95_lower), linetype = "dashed", color = "gray50") +
  
  # Plot the sub-district data points
  geom_point(data = df, aes(x = Positive, y = rate), alpha = 0.6, color = "steelblue") +
  
  # Add the CORRECT Weighted Mean center line
  geom_hline(aes(yintercept = fit.mean), color = "darkgreen", size = 1) +
  
  # Add the INCORRECT Arithmetic Mean center line
  geom_hline(aes(yintercept = simple.mean), color = "red", linetype = "twodash", size = 1) +
  

  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    title = "COVID-19 Case Fatality: Weighted vs Arithmetic Mean",
    subtitle = "Green: Weighted Baseline (Correct) | Red: Arithmetic Baseline (Incorrect)",
    x = "Number of Positive Cases (Sample Size)",
    y = "Mortality Rate"
  ) +
  theme_minimal()

```

## Calculate lower and upper limits for 95% and 99.9% CI

The code chunk below is used to compute the lower and upper limits for 95% confidence interval.

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

## Plotting a static funnel plot

```{r}

p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p
```

## Interactive Funnel Plot: plotly + ggplot2

The funnel plot created using ggplot2 functions can be made interactive with ggplotly() of plotly r package

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```
