---
title: "Hands-on_Ex04"
author: "Alvin Ong Wei Ming"
date: "May 04, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: true #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

# Visualising Distribution

For the purpose of this exercise we will use:

- [ggridges](https://cran.r-project.org/web/packages/ggridges/), a ggplot2 extension specially designed for plotting ridgeline plots,

- [ggdist](https://cran.r-project.org/web/packages/ggdist/index.html), a ggplot2 extension spacially desgin for visualising distribution and uncertainty,
- tidyverse, a family of R packages to meet the modern data science and visual communication needs,

- [ggthemes](https://cran.r-project.org/web/packages/ggthemes/index.html), a ggplot extension that provides the user additional themes, scales, and geoms for the - ggplots package, and

- [colorspace](https://cran.r-project.org/web/packages/colorspace/vignettes/colorspace.html), an R package provides a broad toolbox for selecting individual colors or color palettes, manipulating these colors, and employing them in various kinds of visualisations.

- [BayesFactor](https://cran.r-project.org/web/packages/BayesFactor/index.html), A suite of functions for computing various Bayes factors for simple designs, including contingency tables, one- and two-sample designs, one-way designs, general ANOVA designs, and linear regressio


```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse,BayesFactor)

```

```{r}
exam <- read_csv("data/Exam_data.csv")
head(exam)
```

## Visualising Distribution with Ridgeline Plot
Ridgeline plot (sometimes called Joyplot) is a data visualisation technique for revealing the distribution of a numeric value for several groups. Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.

ggridges package provides two main geom to plot gridgeline plots, they are: geom_ridgeline() and geom_density_ridges(). The former takes height values directly to draw the ridgelines, and the latter first estimates data densities and then draws those using ridgelines.

::: panel-tabset

## Default
```{r}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  )  + scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## With Gradient
```{r}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = " Grades",
                       option = "F") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```
## With Bayes Factor

```{r}
exam <- exam %>%
  mutate(across(CLASS, as.factor))
# 2. Calculate the Bayes Factor 
# This calculates the ratio of the likelihood of the data given the alternate 
# hypothesis (H1) vs. the null hypothesis (H0).
bf_test <- anovaBF(ENGLISH ~ CLASS, data = exam)

# Extract the numeric value of BF10 to use in the plot
bf_val <- extractBF(bf_test)$bf[1]

# 3. Create the Ridgeline Plot with Gradient
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = stat(x))) +
  
  # Create the ridges with a height scale and minimum height threshold
  geom_density_ridges_gradient(
    scale = 3, 
    rel_min_height = 0.01) +
  
  # Apply the Viridis color palette (Option C: Plasma)
  scale_fill_viridis_c(
    name = "Score", 
    option = "F") +
  
  # Clean up the X-axis and remove padding
  scale_x_continuous(
    name = "English Grades", 
    expand = c(0, 0)) +
  
  # Clean up Y-axis and add space at the top for labels
  scale_y_discrete(
    name = NULL, 
    expand = expansion(add = c(0.2, 2.6))) +
  
  # Add the Bayes Factor (BF10) result to the subtitle
  labs(
    title = "Distribution of English Grades by Class",
    subtitle = paste0("Bayes Factor (BF10): ", round(bf_val, 2), 
                      " (Strength of evidence for H1 vs H0)")
  ) +
  
  # Use the specialized ridges theme for a clean look
  theme_ridges()
```
:::


## Mapping the probabilities directly onto color

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```
::: {.callout-important}
## Important

It is important include the argument `calc_ecdf = TRUE` in `stat_density_ridges()`.
:::

## Ridgeline plots with quantile lines

By using [geom_density_ridges_gradient()](https://wilkelab.org/ggridges/reference/geom_ridgeline_gradient.html), we can colour the ridgeline plot by quantile, via the calculated stat(quantile) aesthetic as shown in the figure below.

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```
Instead of using number to define the quantiles, we can also specify quantiles by cut points such as 2.5% and 97.5% tails to colour the ridgeline plot as shown in the figure below.

```{r}

ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()

```
## Visualising Distribution with Raincloud Plot
Raincloud Plot is a data visualisation techniques that produces a half-density to a distribution plot. It gets the name because the density plot is in the shape of a “raincloud”. The raincloud (half-density) plot enhances the traditional box-plot by highlighting multiple modalities (an indicator that groups may exist). The boxplot does not show where densities are clustered, but the raincloud plot does!

::: panel-tabset

## Default
```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               )

```

## Modification

```{r}
# We remove the slab interval by setting .width = 0 and point_colour = NA.

ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA
               )

```
:::

## Adding the boxplot with geom_boxplot()
Next, we will add the second geometry layer using geom_boxplot() of ggplot2. This produces a narrow boxplot. We reduce the width and adjust the opacity.

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)

```

Next, we will add the third geometry layer using stat_dots() of ggdist package. This produces a half-dotplot, which is similar to a histogram that indicates the number of samples (number of dots) in each bin. We select side = “left” to indicate we want it on the left-hand side.
Rainplot looks pretty good here too

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)

```
Flipping it over
```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()

```


```{r}
#| fig-width: 10
#| fig-height: 14
ggplot(exam, aes(x = ENGLISH, y = CLASS, fill = CLASS)) +
  
  # 1. THE "CLOUD" (Half-density curve)
  # justification = -0.2 pushes the density curve up slightly to make room for points
  stat_halfeye(
    adjust = 0.5, 
    justification = -0.1, 
    .width = 0, 
    point_colour = NA
  ) +
  
  # 2. THE BOXPLOT (Quartiles and Median)
  # width = 0.15 keeps it narrow so it doesn't overlap the cloud
  geom_boxplot(
    width = 0.15, 
    outlier.shape = NA, # Hide outliers here so they don't double-print with the dots
    alpha = 0.5
  ) +
  
  # 3. THE "RAIN" (Individual data points)
  # side = "left" puts the dots below the density curve
  # justification = 1.1 ensures the dots start just below the boxplot
  stat_dots(
    side = "left", 
    justification = 1.1, 
    binwidth = NA, # You may need to tweak this value depending on point clustering
    alpha = 0.7
  ) +
  
  # 4. TITLES AND THEME
  labs(
    title = "Distribution of English Grades by Class",
    subtitle = "Bayes Factor (BF10): 4.55",
    x = "English Grades",
    y = NULL
  ) +
  
  # Clean, modern theme
  theme_minimal() +
  theme(
    legend.position = "none", # Hide legend since Y-axis already labels the classes
    panel.grid.minor = element_blank()
  ) +
  scale_fill_viridis_d(option = "plasma", alpha = 0.6) 

```