---
title: "Take Home Exercise 1 - Technical Report"
format:
  html:
    code-fold: true
author: "Alvin Ong Wei Ming"
date: "May 15, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: True #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

## Set Up

Libraries used

-   [ggstatsplot](https://cran.r-project.org/web/packages/ggstatsplot/index.html) Extends ggplot2 by automatically adding information-rich statistical test results (p-values, Bayes factors, effect sizes) directly onto information-dense data visualizations.
-   [ggdist](https://cran.r-project.org/web/packages/ggdist/index.html) Provides a flexible set of specialized ggplot2 geoms and stats designed for visualizing distributions, uncertainty intervals, and Bayesian posterior densities (like half-eye and dotplots).
-   [see](https://cran.r-project.org/web/packages/see/index.html) Serves as the visual engine for the easystats ecosystem, offering clean, publication-ready themes and custom ggplot2 geoms to easily plot complex statistical model outputs.
-   [tidyverse](https://tidyverse.org/) An opinionated collection of core data science packages (like dplyr, ggplot2, and tidyr) designed to work seamlessly together for clean data manipulation, exploration, and visualization.
-   [skimr](https://docs.ropensci.org/skimr/) Provides an elegant, frictionless summary function that produces a beautifully structured console overview of data types, missing value rates, and inline histogram sparklines for a dataset.
-   [naniar](http://naniar.njtierney.com/) Offers a specialized suite of structures and visualizations to look at patterns, percentages, and systemic structures of missing data (NA values) without altering your baseline rows.
-   [scales](https://scales.r-lib.org/) Controls the underlying mathematical transformation, axis formatting, and labeling mechanics of plots, allowing you to easily map numbers to clean currency, percentages, or logarithmic scales.
-   [arrow](https://cran.r-project.org/web/packages/arrow/index.html) Provides an interface to Apache Arrow, enabling ultra-fast read/write data pipelines, out-of-memory computations, and seamless querying of massive parquet file portfolios.
-   [patchwork](https://patchwork.data-imaginist.com/) Makes combining multiple separate ggplot2 charts into cohesive, multi-panel layouts or slide grids as simple as using basic arithmetic operators like + and /.
-   [GGally](https://cran.r-project.org/web/packages/GGally/index.html) Extends ggplot2 by combining multiple columns into compound visual layouts, such as pairwise correlation matrix grids (pairs plots) and parallel coordinate plots.
-   [glue](https://tidyverse.r-universe.dev/articles/glue/glue.html) Replicates the smooth developer experience of Python’s f-strings in R, allowing you to cleanly evaluate expressions and parse variables directly inside text strings using curly braces {}.
-   [ggpubr](https://cran.r-project.org/web/packages/ggpubr/index.html) Wraps complex ggplot2 commands into simpler functions optimized for creating publication-ready, clean statistical charts for research papers and academic panels.
-   [treemapify](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) Adds custom tiles and text-fitting geoms to ggplot2 to easily build nested, variable-area treemaps for visualizing categorical concentration and financial value distributions.
-   [vcd](https://cran.r-project.org/web/packages/vcd/index.html) Provides a highly structured suite of statistical tools and custom visual plots (like mosaic and association plots) designed specifically to evaluate contingency tables and categorical data associations.

### Load Packages

```{r}
#| code-fold: false
pacman::p_load(ggstatsplot, 
               ggdist,
               see,
               tidyverse,
               skimr,
               naniar,
               scales,
               arrow,
               patchwork,
               GGally,
               glue,
               ggpubr,
               treemapify,
               vcd)
```

### Colour used

```{r}
color_palette <- tibble::tribble(
  ~Context,                            ~Hex_Code,
  "Comparison - Before State",         "#f8766d",
  "Comparison - After State",          "#00bfc4",
  "Background Secondary Tone",         "#4d4d4d", # 'grey30' hex equivalent
  "Financial Metric - Premiums",       "#2ca02c",
  "Financial Metric - Claims",         "#d9534f",
  "Reference - Average Dash Lines",    "#666666", # 'gray40' hex equivalent
  "Premium Line Chart Accent",         "#D4AF37",
  "Green Gradient Baseline (Low)",     "#A2E1A6",
  "Green Gradient Peak (High)",        "#137333",
  "Red Gradient Baseline (Low)",       "#3A1414",
  "Red Gradient Peak (High)",          "#E57373",
  "Sparsity - Claim Filed (1)",        "#E46726",
  "Sparsity - No Claim (0)",           "#F0F0F0",
  "Divergence Scale - Low Anchor",     "#2c7bb6",
  "Divergence Scale - Mid Anchor",     "#ffffbf",
  "Divergence Scale - High Anchor",    "#d7191c",
  "Categorical Association - Mid",     "#fdae61"
)

color_preview_table <- color_palette %>%
  mutate(
    Clean_Hex = str_remove(Hex_Code, "#"),
    Preview_Link = glue("[View {Hex_Code} Visual](https://www.color-hex.com/color/{Clean_Hex})")
  ) %>%
  select(Context, `Hex Code` = Hex_Code, `Live Link` = Preview_Link)

knitr::kable(color_preview_table, format = "markdown")
```

### Read in Data


```{r}
#| code-fold: false
df <-read.csv2("Dataset of motor insurance portfolio.csv")
```

## Data Overview

### See overall and missing values.

```{r}

#notice that fuel type has 1287 missing values
# vehicle type has 513 value
# vehicle_age has 2 missing value
# age_driving_license has 2 missing values
skim(df)
```

### See Data types'

```{r}
str(df)
```

### Total Exposure max value

motor insurance happen to have only max 1 year contract.

```{r}
max(df$total_exposure)
```

### Identify if the missingness has a pattern.

An Upset-plot is generated to identify if the missingness of value has a pattern. We observe that the missing values for vehicle values and fuel_type are likely independent. Whereas vehicle and age are missing together but there are only 2 rows affected by it. If we are using vehicle values and fuel_type [Missing at Random](https://medium.com/@adnan.mazraeh1993/type-of-missing-data-c52b7788da52) seems to be in play here

```{r}
gg_miss_upset(df, nsets = 4)

```

### Distribution of vehicle value and fuel type

::: panel-tabset
### Vehicle Value Distribution

By Default ggplot has removed the null values here and a right skew is observed for this column.

```{r}
ggplot(data=df, 
             aes(x = as.integer(vehicle_value))) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +

  scale_x_continuous(labels = label_comma()) +
  theme_gray() +
  ggtitle("Distribution of Vehicle Value") +
labs(x = "Vehicle Value", y = "Count")
```

### Fuel Type Distribution

```{r}

df_clean <- df %>%
  mutate(
    fuel_type = case_when(
      fuel_type == 'D' ~ "Diesel",
      fuel_type == 'G' ~ "Gasoline",
      is.na(fuel_type) ~ "Unknown", # Label the 36% missing values explicitly
      TRUE             ~ "Others"
    ),
    fuel_type = factor(fuel_type, levels = c("Gasoline", "Diesel", "Unknown", "Others"))
  )

# 2. Build the Bar Chart Layer (p1)
p1 <- ggplot(data = df_clean, aes(x = fuel_type)) +
  geom_bar(fill = "grey70", color = "grey25", alpha = 0.8) +
  theme_minimal() +
  labs(
    title = "Frequency Distribution of Fuel Types",
    subtitle = "Explicitly highlighting missing data as 'Unknown'",
    x = "Fuel Type Classification", 
    y = "Count",
    fill = "Fuel Type"
  )

# 3. Build the Numerical Missingness Audit Data Frame
missing_audit_data <- df %>%
  summarise(
    `Total Missing Rows` = format(sum(is.na(fuel_type)), big.mark = ","),
    `Missing Percentage` = paste0(round(mean(is.na(fuel_type)) * 100, 2), "%")
  )

# 4. Turn the data frame into a graphical table layer (p2)
p2 <- ggtexttable(missing_audit_data, rows = NULL)

# 5. Stitch them together vertically using patchwork
p1 / p2 + patchwork::plot_layout(heights = c(4, 1))
```

```{r}
df %>%
  summarise(
    Total_Missing_Rows = sum(is.na(fuel_type)),
    Percentage_Missing = mean(is.na(fuel_type)) * 100 # If this is > 5%, be careful with the mode!
  )
```
:::

## Data Cleaning

### Clean up of data, data type conversion and new variable creation

-   Convert all categorical values with acronym into readable label
-   Remove the 2 rows of data where vehicle age and age driving license is null
-   Impute Median to vehicle age and mode to fuel type
-   We also created indicator called underwriting profit to identify which portfolios are losses.

```{r}

# Create function to calculate mode
calc_mode <- function(x) {
  distinct_values <- unique(x[!is.na(x)])
  distinct_values[which.max(tabulate(match(x, distinct_values)))]
}

# Pre-calculate the mode safely outside the pipeline to unblock the compiler
fuel_mode <- calc_mode(df$fuel_type)

df_transformed <- df %>%
  mutate(
    # All the factor like variables
    policy_type = case_when(
      policy_type == "TP"     ~ "Basic Third-Party",
      policy_type == "TPG"    ~ "Third-Party with Glass Cov",
      policy_type == "CC"     ~ "Third-Party with two or more Cov",
      policy_type == "COMP_E" ~ "Comprehensive with Excess",
      policy_type == "COMP_N" ~ "Comprehensive without Excess",
      TRUE                    ~ "Others"
    ),
    policy_status = case_when(
      policy_status == 'A' ~ "Active",
      policy_status == 'C' ~ "Cancelled",
      TRUE                 ~ "Others"
    ),
    business_type = case_when(
      business_type == 'NB' ~ "New_Business",
      business_type == 'P' ~ "Portfolio",
      TRUE                 ~ "Others"
    ),
    payment_frequency = case_when(
      payment_frequency == 'A' ~ "Annual",
      payment_frequency == 'S' ~ "Semi-Annual",
      payment_frequency == 'Q' ~ "Quarterly",
      TRUE                     ~ "Others"
    ),
    bonus_score = case_when(
      bonus_score == 'G' ~ "Positive",
      bonus_score == 'N' ~ "Neutral",
      bonus_score == 'B' ~ "Negative",
      TRUE                 ~ "Others"
    ),
    fuel_type = case_when(
      fuel_type == 'D' ~ "Diesel",
      fuel_type == 'G' ~ "Gasoline",
      is.na(fuel_type) ~ NA_character_,
      TRUE                 ~ "Others"
    ),
    municipality_type = case_when(
      municipality_type == 'I' ~ "Inland",
      municipality_type == 'C' ~ "Coastal",
      municipality_type == 'IS' ~ "Islands",
      TRUE                     ~ "Others"
    ),
    circulation_area = case_when(
      circulation_area == 'U' ~ "Urban",
      circulation_area == 'R' ~ "Rural",
      TRUE                    ~ "Others"
    ),

    # These are all the numeric conversions
    vehicle_value = as.numeric(vehicle_value),
    power_to_weight_ratio = as.numeric(power_to_weight_ratio),
    total_premium = as.numeric(total_premium),
    liability_premium = as.numeric(liability_premium),
    property_damage_premium = as.numeric(property_damage_premium),
    theft_premium = as.numeric(theft_premium),
    fire_premium = as.numeric(fire_premium),
    glass_premium = as.numeric(glass_premium),
    legal_protection_premium = as.numeric(legal_protection_premium),
    occupants_premium = as.numeric(occupants_premium),
    total_incurred = as.numeric(total_incurred),
    liability_incurred = as.numeric(liability_incurred),
    liability_property_incurred = as.numeric(liability_property_incurred),
    liability_injury_incurred = as.numeric(liability_injury_incurred),
    property_incurred = as.numeric(property_incurred),
    theft_incurred = as.numeric(theft_incurred),
    fire_incurred = as.numeric(fire_incurred),
    glass_incurred = as.numeric(glass_incurred),
    legal_protection_incurred = as.numeric(legal_protection_incurred),
    occupants_incurred = as.numeric(occupants_incurred),
    total_exposure = as.numeric(total_exposure),
    liability_exposure = as.numeric(liability_exposure),

    # New variables created
    underwriting_profit = total_premium - total_incurred
  ) %>%
  mutate(
    policy_type = factor(policy_type, levels = c(
      "Basic Third-Party",
      "Third-Party with Glass Cov",
      "Third-Party with two or more Cov",
      "Comprehensive with Excess",
      "Comprehensive without Excess",
      "Others"
    )),
    # Directly reference the column in median calculation to be safe
    vehicle_value = replace_na(vehicle_value, median(vehicle_value, na.rm = TRUE)),
    policy_status = as.factor(policy_status),
    business_type = as.factor(business_type),
    payment_frequency = as.factor(payment_frequency),
    # Use the static pre-calculated scalar value here
    fuel_type = as.factor(replace_na(fuel_type, fuel_mode)),
    municipality_type = as.factor(municipality_type),
    circulation_area = as.factor(circulation_area),
    bonus_score = as.factor(bonus_score)
  ) %>%
  filter(!is.na(vehicle_age))

```

### For columns imputed check distribution before and after

```{r}

# 2. Re-create your comparative dataset (Long format)
df_before <- df %>%
  select(vehicle_value) %>%
  mutate(vehicle_value = as.numeric(vehicle_value), status = "Before")

df_after <- df_transformed %>%
  select(vehicle_value) %>%
  mutate(status = "After")

df_compare <- bind_rows(df_before, df_after) %>% filter(!is.na(vehicle_value))

# 2. Generate the side-by-side comparative violin plot safely
p1 <- ggplot(df_compare, aes(x = "Vehicle Value Profile", y = vehicle_value, fill = status)) +
  # Layer 1: Draw Left Half (Before Imputation)
  see::geom_violinhalf(
    data = filter(df_compare, status == "Before"),
    flip = FALSE, 
    alpha = 0.6, 
    color = "grey25"
  ) + 
  # Layer 2: Draw Right Half (After Imputation)
  see::geom_violinhalf(
    data = filter(df_compare, status == "After"),
    flip = TRUE,   # Only flip this specific layer
    alpha = 0.6, 
    color = "grey25"
  ) +
  scale_fill_manual(values = c("Before" = "#f8766d", "After" = "#00bfc4")) +
  scale_y_continuous(labels = label_comma()) +
  theme_minimal() +
  labs(
    title = "Imputation Audit: Vehicle Value Distribution",
    subtitle = "Comparing raw data density against median-imputed data",
    x = "", 
    y = "Vehicle Value ($)",
    fill = "Status"
  )

summary_table_data <- df_compare %>%
  group_by(status) %>%
  summarise(
    Total_Rows = comma(n()),
    Mean_Value = comma(mean(vehicle_value), accuracy = 1),
    Median_Val = comma(median(vehicle_value), accuracy = 1),
    Std_Dev    = comma(sd(vehicle_value), accuracy = 1)
  ) %>%
  rename(Status = status, `Total Rows` = Total_Rows, `Mean ($)` = Mean_Value, `Median ($)` = Median_Val, `SD ($)` = Std_Dev)

# 4. Create the Graphic Table Layer (p2)
p2 <- ggtexttable(summary_table_data, rows = NULL)

# 5. Use Patchwork to stitch them vertically and set heights
# The plot takes 80% of the vertical space, the table takes 20%
p1 / p2 + patchwork::plot_layout(heights = c(4, 1))
```

```{r}
cat("--- FUEL PROPORTIONS PROFILE before imputation ---\n")

df %>%
  count(fuel_type) %>%
  mutate(Percentage_Share = paste0(round((n / sum(n)) * 100, 2), "%")) %>%
  print()



cat("--- FUEL PROPORTIONS PROFILE after imputation---\n")

df_transformed %>%
  count(fuel_type) %>%
  mutate(Percentage_Share = paste0(round((n / sum(n)) * 100, 2), "%")) %>%
  print()
```

## Clean Data writeout

-   We will write out the data to parquet format to ensure the metadata and datatype is preserved
-   the arrow package is used.
-   This copy can be saved in an SQL server for safekeeping for future reports and reference.

```{r}
#| code-fold: false
#| eval: false
write_parquet(df_transformed, sink = "df_transformed.parquet")
```

## Re-read the Transformed Data

```{r}
#| code-fold: false
df_transformed =  read_parquet("df_transformed.parquet")
# dplyr::glimpse(df_transformed)
```

## Analysis

### 1. Financial health of the business

-   Chart A
    -   We want to display the overall premiums earned and claims paid out to understand the trending and year on year change for each year.
-   Chart B
    -   We want to show percentage change in net profitability from year to year taking into account this years profit - previous year profit / absolute of previous year profit
-   Chart C
    -   We wish to understand for each unit of exposure, how much dollar are we expecting, is the trade off increasing or decreasing

```{r}
#| fig-width: 16
#| fig-height: 8

yearly_summary <- df_transformed %>%
  group_by(year) %>%
  summarize(
    sum_premium = sum(total_premium, na.rm = TRUE),
    sum_incurred = sum(total_incurred, na.rm = TRUE),
    underwriting_profit = sum_premium - sum_incurred,
    avg_rev_per_exposure = sum_premium / sum(total_exposure, na.rm = TRUE)
  ) %>%
  arrange(year) %>%
  # Calculate YoY changes
  mutate(
    premium_yoy = (sum_premium - lag(sum_premium)) / lag(sum_premium) * 100,
    incurred_yoy = (sum_incurred - lag(sum_incurred)) / lag(sum_incurred) * 100
  )

# =========================================================================
# STEP 2: Pivot & Format Labels for Plot 1 (Grouped Bars)
# =========================================================================
p1_data <- yearly_summary %>%
  select(year, `Total Premium` = sum_premium, `Total Incurred Claims` = sum_incurred, premium_yoy, incurred_yoy) %>%
  pivot_longer(
    cols = c(`Total Premium`, `Total Incurred Claims`), 
    names_to = "Metric", 
    values_to = "Amount"
  ) %>%
  # Assign the correct YoY percentage to the correct metric row
  mutate(
    yoy_val = if_else(Metric == "Total Premium", premium_yoy, incurred_yoy),
    # Format text: If it's the first year, just show amount. Otherwise, show amount + YoY %
    label_text = case_when(
      is.na(yoy_val) ~ label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount),
      yoy_val >= 0   ~ paste0(label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount), "\n(+", round(yoy_val, 1), "%)"),
      TRUE           ~ paste0(label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount), "\n(", round(yoy_val, 1), "%)")
    )
  )

# =========================================================================
# STEP 3: Build Plot 1 - Grouped Bar Chart with Value & % Labels
# =========================================================================
# Explicitly defining bar positioning width so the text alignments align flawlessly
bar_width <- 0.6
dodge_width <- 0.7

p1 <- ggplot(p1_data, aes(x = factor(year), y = Amount, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(width = dodge_width), width = bar_width) +
  
  # Add dynamic text labels on top of bars
  geom_text(
    aes(label = label_text),
    position = position_dodge(width = dodge_width),
    vjust = -0.25,       
    size = 3,            
    fontface = "bold",
    lineheight = 0.85    
  ) +
  
  scale_fill_manual(values = c("Total Premium" = "#2ca02c", "Total Incurred Claims" = "#d9534f")) +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0, 0.15)) # Adds 15% headroom at the top so labels don't clip
  ) +
  labs(
    title = "A: Premium Inflow vs. Claims Outflow",
    x = NULL, 
    y = "Total Financial Amount (Y-O-Y % Increase)", 
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 4: Build Plot 2 - Net Underwriting Profit Trend with Labels
# =========================================================================
yearly_summary <- yearly_summary %>%
  mutate(
    profit_label = label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(underwriting_profit)
  )

p2 <- ggplot(yearly_summary, aes(x = factor(year), y = underwriting_profit, fill = underwriting_profit > 0)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.9) +
  geom_text(
    aes(label = profit_label, vjust = if_else(underwriting_profit >= 0, -0.4, 1.4)),
    size = 3,
    fontface = "bold"
  ) +
  scale_fill_manual(values = c("TRUE" = "#2ca02c", "FALSE" = "#d9534f"), guide = "none") +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0.15, 0.15))
  ) +
  labs(
    title = "B: Net Underwriting Profit Trend",
    subtitle = "Green = Profitable | Red = Underwriting Loss",
    x = NULL, 
    y = "Net Profit"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 9, color = "gray30"),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 5: Build Plot 3 - Average Revenue Per Unit of Exposure
# =========================================================================
p3 <- ggplot(yearly_summary, aes(x = factor(year), y = avg_rev_per_exposure, group = 1)) +
  geom_line(color = "#D4AF37", linewidth = 1.2) +
  geom_point(color = "#1c3b57", size = 3) +
  geom_text(
    aes(label = paste0("$", round(avg_rev_per_exposure, 1))), 
    vjust = -1, 
    size = 3.5, 
    fontface = "bold"
  ) +
  scale_y_continuous(
    labels = label_number(prefix = "$"),
    expand = expansion(mult = c(0.1, 0.2))
  ) +
  labs(
    title = "C: Average Premium Revenue per Unit of Exposure Year",
    x = "Year", 
    y = "Avg Premium / Exposure"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 6: Assemble Framework Dashboard Layout and Print
# =========================================================================
dashboard <- (p1 | p2) / p3 + 
  plot_annotation(
    title = "Overall Financial Health",
    subtitle = "Year-on-Year metrics tracking revenue growth, claims impact, and net margin",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, margin = margin(b = 5)),
      plot.subtitle = element_text(color = "gray25", size = 11, margin = margin(b = 15))
    )
  )

# Preview the clean dashboard
print(dashboard)
```

### 2.Product profitability from 2022 to 2024

In this section, we want to see the overall view of the motor insurance view. How much of the portfolio belongs to each category of product.

```{r}
#| fig-width: 18
#| fig-height: 8
#| dpi: 150


premium_summary <- df_transformed %>%
  group_by(policy_type) %>%
  summarize(total_premium_sum = sum(total_premium, na.rm = TRUE))


incurred_summary <- df_transformed %>%
  group_by(policy_type) %>%
  summarize(total_incurred_sum = sum(total_incurred, na.rm = TRUE))

# Merge and calculate cumulative metrics based on Premium Rank
pareto_combined <- premium_summary %>%
  left_join(incurred_summary, by = "policy_type") %>%
  arrange(desc(total_premium_sum)) %>%
  mutate(
    policy_type = factor(policy_type, levels = policy_type),
    
    # Premium Tail Metrics
    cum_sum_prem = cumsum(total_premium_sum),
    total_sum_prem = sum(total_premium_sum),
    cum_pct_prem = (cum_sum_prem / total_sum_prem) * 100,
    remaining_pct_prem = 100 - lag(cum_pct_prem, default = 0),
    
    # Claims Tail Metrics
    cum_sum_inc = cumsum(total_incurred_sum),
    total_sum_inc = sum(total_incurred_sum),
    cum_pct_inc = (cum_sum_inc / total_sum_inc) * 100,
    remaining_pct_inc = 100 - lag(cum_pct_inc, default = 0)
  )

# Master anchors for dual-axis transformation
total_premium_total <- sum(pareto_combined$total_premium_sum)
max_y_limit <- total_premium_total * 1.1

# Pivot long for side-by-side dodge bars
bar_data <- pareto_combined %>%
  select(policy_type, Premium = total_premium_sum, Claims = total_incurred_sum) %>%
  pivot_longer(cols = c(Premium, Claims), names_to = "Metric", values_to = "Amount") %>%
  mutate(Metric = factor(Metric, levels = c("Premium", "Claims")))

# --- 2. GENERATE OVERLAYED PLOT ---
ggplot() +
  # Side-by-side grouped bars for Premium vs Claims
  geom_bar(data = bar_data, aes(x = policy_type, y = Amount, fill = Metric), 
           stat = "identity", position = position_dodge(width = 0.8), alpha = 0.85, width = 0.7) +
  
  # Gold Line: Remaining Revenue Tail
  geom_line(data = pareto_combined, aes(x = as.numeric(policy_type) - 0.2, 
                                        y = remaining_pct_prem * (total_premium_total / 100), group = 1, color = "Remaining Revenue Tail"), 
            linewidth = 1.2) +
  geom_point(data = pareto_combined, aes(x = as.numeric(policy_type) - 0.2, 
                                         y = remaining_pct_prem * (total_premium_total / 100)), color = "#D4AF37", size = 2.5) +
  
  # Dual Axis Configuration
  scale_y_continuous(
    name = "Financial Amount ($)",
    limits = c(0, max_y_limit),
    breaks = seq(0, max_y_limit, by = 10000000),
    labels = label_comma(),
    sec.axis = sec_axis(
      trans = ~ . / total_premium_total * 100,
      name = "Remaining Portfolio Percentage (%)",
      breaks = seq(0, 100, by = 10)
    )
  ) +
  
  # Styling and Legends
  scale_fill_manual(values = c("Premium" = "#2ca02c", "Claims" = "#d9534f")) +
  scale_color_manual(values = c("Remaining Revenue Tail" = "#D4AF37", "Remaining Claims Tail" = "grey25")) +
  
  labs(
    title = "Portfolio Premium & Incurred Claims",
    subtitle = "Bars show Total Premium (Green)/ Claims (Red).\n Gold Line shows the propotion of portfolio.",
    x = "Policy Type",
    fill = "Financial Metrics",
    color = "Portfolio Tails"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  )
```

::: callout-tip
## Findings

-   Comprehensive without Excess Insurance has very high claims , rendering the product to be either break even or barely profitable
-   Third Party with glass coverage insurance also has very high claim amount and thin margins
-   Basic Third party insurance has razor thin margins
:::

### 3. Premiums from 2022 to 2024

-   This is to perform confirmatory data analysis that there is a change in distribution amongst the various product over the years

-   Where the Null Hypothesis: The distribution of total premium is identical across all three years Alternative Hypothesis: The distribution of total premium is not identical across all three years; at least one year's premium distribution stochastically dominates

-   We also observed that there are extreme values in total premiums, hence a log is applied to the y axis to help observe the full range of premiums more easily

### Create function

```{r}
plot_policy_distribution <- function(data, target_policy, target_metric = "total_premium") {
  
  # 1. Dynamically filter and prepare data based on function arguments
  # We use .data[[target_metric]] to allow passing the column name as a string
  plot_data <- data %>%
    filter(
      policy_type == target_policy,
      .data[[target_metric]] > 0
    ) %>%
    mutate(year = factor(year))
  
  # Determine dynamic text for labels
  metric_label <- gsub("_", " ", target_metric) %>% tools::toTitleCase()
  
  # 2. Safe Statistical Check
  unique_years <- unique(plot_data$year)
  
  if (length(unique_years) > 1) {
    # Dynamically build formula: target_metric ~ year
    formula_obj <- as.formula(paste(target_metric, "~ year"))
    stat_test   <- kruskal.test(formula_obj, data = plot_data)
    
    stat_label  <- paste0(
      "Kruskal-Wallis chi-squared = ", round(stat_test$statistic, 2),
      ", df = ", stat_test$parameter,
      ", p-value = ", scales::scientific(stat_test$p.value, digits = 3)
    )
  } else if (length(unique_years) == 1) {
    stat_label <- paste0("Kruskal-Wallis: Insufficient groups (Only ", unique_years, " data available)")
  } else {
    # If string matching yields 0 rows, print warning and stop to avoid ggplot crash
    warning(paste0("No data found for policy type: '", target_policy, "' with ", metric_label, " > 0."))
    return(NULL)
  }
  
  # 3. Generate the Plot
  p <- ggplot(plot_data, aes(x = year, y = .data[[target_metric]], fill = year)) +
    
    # The Half-Eye distribution view
    stat_halfeye(
      adjust = 0.5,
      width = 0.6,
      .width = c(0.66, 0.95),
      point_color = "black",
      interval_color = "black",
      alpha = 0.7
    ) +
    
    # Scale and formatting
    scale_y_log10(labels = scales::label_comma(prefix = "$")) +
    scale_fill_brewer(palette = "Blues", guide = "none") +
    
    # Informative labels that update dynamically based on your inputs
    labs(
      title = paste0(metric_label, " Distribution across Years"),
      subtitle = paste0(target_policy, "\n", stat_label),
      x = "Year",
      y = paste0(metric_label, " (Log Scale)")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(face = "bold")
    ) +
  coord_flip() 
  return(p)
}


plot_pairwise_matrix <- function(data, target_policy, target_metric = "total_premium") {
  
  # 1. Filter data based on input arguments
  plot_data <- data %>%
    filter(
      policy_type == target_policy,
      .data[[target_metric]] > 0
    ) %>%
    mutate(year = factor(year))
  
  # Clean up the metric name for beautiful axis/title labels
  metric_label <- gsub("_", " ", target_metric) %>% tools::toTitleCase()
  
  # 2. Check if pairwise testing is mathematically possible
  if (length(unique(plot_data$year)) < 2) {
    warning(paste0("Skipping matrix: '", target_policy, "' does not have enough distinct years with data."))
    return(NULL)
  }
  
  # 3. Execute the Pairwise Wilcoxon Test with Holm Adjustment
  wt_result <- pairwise.wilcox.test(
    x = plot_data[[target_metric]], 
    g = plot_data$year,
    p.adjust.method = "holm"
  )
  
  p_matrix <- wt_result$p.value
  
  # Handle edge case where the matrix is completely empty/NULL
  if (all(is.na(p_matrix))) {
    warning(paste0("No valid variance found to compute pairwise shifts for '", target_policy, "'."))
    return(NULL)
  }
  
  # 4. Transform matrix to a tidy long-format dataframe
  p_df_long <- as.data.frame(p_matrix) %>%
    mutate(Group_B = rownames(.)) %>%
    pivot_longer(
      cols = -Group_B, 
      names_to = "Group_A", 
      values_to = "p_val"
    ) %>%
    filter(!is.na(p_val)) %>%
    mutate(
      # Ensure factor matching for explicit manual scale mapping
      is_significant = factor(p_val < 0.05, levels = c("TRUE", "FALSE")),
      p_label = if_else(p_val < 0.001, "p < 0.001", paste0("p = ", round(p_val, 3)))
    )
  
  # 5. Build the Dynamic Matrix Plot
  p <- ggplot(p_df_long, aes(x = Group_A, y = Group_B, fill = is_significant)) +
    geom_tile(color = "white", linewidth = 1.5) +
    geom_text(aes(label = p_label), color = "white", fontface = "bold", size = 4.5) +
    
    # Red/Coral for significant shifts, neutral grey for stagnant periods
    scale_fill_manual(
      values = c("TRUE" = "#2ca02c", "FALSE" = "#7f8c8d"),
      labels = c("TRUE" = "Significant Shift (p < 0.05)", "FALSE" = "No Stat. Difference"),
      name = "Holm Conclusion",
      drop = FALSE 
    ) + 
    
    # FORCE the legend preview boxes to show their true assigned colors manually
    guides(
      fill = guide_legend(
        override.aes = list(fill = c("#2ca02c", "#7f8c8d"))
      )
    ) + 
    
    labs(
      title = paste0("Pairwise Post-Hoc Shifts Matrix (", metric_label, ")"),
      subtitle = paste0(target_policy, "\nWilcoxon Rank-Sum Test (Holm-Adjusted Method)"),
      x = "Baseline Year (Group A)",
      y = "Comparison Year (Group B)"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray30", face = "italic", size = 10, margin = margin(b = 10)),
      axis.text = element_text(face = "bold", size = 11),
      panel.grid = element_blank(),
      legend.position = "bottom",
      legend.title = element_text(face = "bold")
    )
  
  return(p)
}
```

::: panel-tabset
### Third-Party with two or more cov

```{r}
plot_policy_distribution(df_transformed, "Third-Party with two or more Cov")

```

```{r}
plot_pairwise_matrix(df_transformed, "Third-Party with two or more Cov")
```

### Comprehensive with Excess

```{r}
plot_policy_distribution(df_transformed, "Comprehensive with Excess")
```

```{r}
plot_pairwise_matrix(df_transformed, "Comprehensive with Excess")

```

### Comprehensive without Excess

```{r}
plot_policy_distribution(df_transformed, "Comprehensive without Excess")
```

```{r}
plot_pairwise_matrix(df_transformed, "Comprehensive without Excess")

```

### Third-Party with Glass Cov

```{r}
plot_policy_distribution(df_transformed, "Third-Party with Glass Cov")
```

```{r}
plot_pairwise_matrix(df_transformed, "Third-Party with Glass Cov")
```

### Basic Third-Party

```{r}
plot_policy_distribution(df_transformed, "Basic Third-Party")
```

```{r}

plot_pairwise_matrix(df_transformed, "Basic Third-Party")

```
:::

## 4. Claims from 2022 to 2024

Function created to ease the coding below. Instead of blue we will use red for claims for the violin plot below. We also perform a log in the Y axis to observe the full range of values.

```{r}
plot_claims_distribution <- function(data, target_policy, target_metric = "total_incurred") {
  
  # 1. Dynamically filter and prepare data based on function arguments
  # We use .data[[target_metric]] to allow passing the column name as a string
  plot_data <- data %>%
    filter(
      policy_type == target_policy,
      .data[[target_metric]] > 0
    ) %>%
    mutate(year = factor(year))
  
  # Determine dynamic text for labels
  metric_label <- gsub("_", " ", target_metric) %>% tools::toTitleCase()
  
  # 2. Safe Statistical Check
  unique_years <- unique(plot_data$year)
  
  if (length(unique_years) > 1) {
    # Dynamically build formula: target_metric ~ year
    formula_obj <- as.formula(paste(target_metric, "~ year"))
    stat_test   <- kruskal.test(formula_obj, data = plot_data)
    
    stat_label  <- paste0(
      "Kruskal-Wallis chi-squared = ", round(stat_test$statistic, 2),
      ", df = ", stat_test$parameter,
      ", p-value = ", scales::scientific(stat_test$p.value, digits = 3)
    )
  } else if (length(unique_years) == 1) {
    stat_label <- paste0("Kruskal-Wallis: Insufficient groups (Only ", unique_years, " data available)")
  } else {
    # If string matching yields 0 rows, print warning and stop to avoid ggplot crash
    warning(paste0("No data found for policy type: '", target_policy, "' with ", metric_label, " > 0."))
    return(NULL)
  }
  
  # 3. Generate the Plot
  p <- ggplot(plot_data, aes(x = year, y = .data[[target_metric]], fill = year)) +
    
    stat_halfeye(
      width = 0.6,
      .width = c(0.66, 0.95),
      point_color = "black",
      interval_color = "black",
      alpha = 0.7
    )  + 
  scale_y_log10(
    limits = c(0.10, 500000), 
    labels = scales::label_dollar(),
    breaks = c(0.10, 1, 10, 100, 1000, 10000, 100000, 500000)
  ) +
    scale_fill_brewer(palette = "Reds", guide = "none") +

    labs(
      title = paste0(metric_label, " Distribution across Years"),
      subtitle = paste0(target_policy, "\n", stat_label),
      x = "Year",
  y = paste0(metric_label, " (Log Scale)")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(face = "bold")
    ) +
  coord_flip() 
  return(p)
}

```

::: panel-tabset
### Third-Party with two or more Cov

```{r}
plot_claims_distribution(df_transformed,"Third-Party with two or more Cov")
```

```{r}

plot_pairwise_matrix(df_transformed, "Third-Party with two or more Cov","total_incurred")

```

### Comprehensive with Excess

```{r}
plot_claims_distribution(df_transformed,"Comprehensive with Excess")
```

```{r}

plot_pairwise_matrix(df_transformed, "Comprehensive with Excess","total_incurred")

```

### Comprehensive without Excess

```{r}
plot_claims_distribution(df_transformed,"Comprehensive without Excess")
```

```{r}

plot_pairwise_matrix(df_transformed, "Comprehensive without Excess","total_incurred")

```

### Third-Party with Glass Cov

```{r}
plot_claims_distribution(df_transformed,"Third-Party with Glass Cov")
```

```{r}

plot_pairwise_matrix(df_transformed, "Third-Party with Glass Cov","total_incurred")

```

### Basic Third-Party

```{r}
plot_claims_distribution(df_transformed,"Basic Third-Party")
```

```{r}

plot_pairwise_matrix(df_transformed, "Basic Third-Party","total_incurred")

```
:::

### Additional charts for stakeholders (Premiums and claims from 2022 to 2024)

This is to emphasise and focus on the peak of for Third-Party with Glass Coverage

```{r}

ggplot(df_transformed %>% filter(policy_type == "Third-Party with Glass Cov"), 
       aes(x = factor(year), y = total_premium, fill = factor(year))) +
  
  # Keeps your original smooth halfeye
  ggdist::stat_halfeye(
    adjust = 0.5,
    width = 0.6,
    .width = c(0.66, 0.95),
    point_color = "black",
    interval_color = "black",
    alpha = 0.7
  ) +
  
  scale_y_continuous(labels = scales::label_comma(prefix = "$")) +

  coord_cartesian(ylim = c(0, 200)) + 
  
  scale_fill_brewer(palette = "Blues", guide = "none") +
  
  labs(
    title = "Total Premium Distribution across Years",
    subtitle = "Third-Party with Glass Cov",
    x = "Year",
    y = "Total Premium"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
    panel.grid.minor = element_blank(),
    axis.text = element_text(face = "bold")
  ) +
  
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() 
```

### Observation from Customer Plans

::: callout-note
! important
Note: The scale is in log, but the x axis is in actual dollars

:::

⚠️Third-Party with two or more Cov:

-   Total Premium Distribution change is associated with more lower priced premiums, although the median remain relatively similar.

-   Median of claims stayed the same from 2022 to 2024 but the number of $1000  claims is more concentrated. There are also more variation of claims below the 25th percentile of all claims

✅Comprehensive with excess:

-   Distribution change seems to be associated with more lower priced premiums, although the median remain relatively similar.

-   On a year on year basis, there is no big change in the claims and its stable

⚠️Comprehensive without excess:

-   Pricing for this category seems to have move to a higher tier as compared to 2023 and 2022 as more motor insurance plans has shifted to a higher premium tier. This is observed from the two peaks in 2024

-   For claims, there are two distinct peaks in the distribution, and there is no change in distribution between 2023 and 2024.

-   Claim and pricing display similar patterns and no drastic movement is observed

::: callout-note
## Key Underwriting Insight

Although on an aggregate level, Comprehensive insurance without excess is barely making a profit. A change seems to have taken place between 2023 and 2024. The diverging two peaks that form in 2024 appears to be two different sub groups of customers. But most importantly, the change has taken effect and appears to be working.
:::

✅Third Party with Glass Cov

-   Although there is change in distribution when comparing the 3 years. The year 2022 and 2024 ($p-value = 0.043$) is very close to the alpha value tolerance level of 0.05. This could mean that the distribution is nearing the distribution from 2024 to 2022.

-   Claims stayed relative stable and predictable.

⚠️Basic Third Party Plan:

-   The Interquartile Range (IQR) for premiums compressed significantly in 2024 compared to prior years. Notably, the 2024 violin plot exhibits a clear bimodal (split) peak, indicating emerging sub-groups or distinct risk segmentation within the product line. Pairwise testing indicates that while overall historical changes are decisive , the shift in mean premiums between 2023 and 2024 was not statistically significant ($p = 0.062$).

-   Claim pattern have change as well from the bulge in claims around the medium claims in 2024 as compared to 2022 and 2023. IQR now is longer and less predictable.

- This also means that this portfolio will need to be revisited as more claims hit the median and the overall range of claims are wider and less predictable 

## 5. Overall Age Groups that has the most claims

```{r}
heatmap_data <- df_transformed %>%
  # Create logical bins for your continuous variables
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  # Group by the bins to isolate each intersection pocket
  group_by(age_bin, v_age_bin) %>%
  # Calculate aggregate metrics per cell
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    # Calculate the Claim Rate (Frequency per policy/exposure)
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  # Optional: Filter out empty or extremely low-volume cells to remove noise
  filter(total_policies > 10)

# 2. Plot the binned matrix
ggplot(heatmap_data, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  # Use a color scale that makes high values (the spike) pop out vividly
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  # Add the actual text values inside the tiles for quick confirmation
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3.5, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Identifying localized risk spikes across driver and vehicle age pockets",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 11),
    panel.grid = element_blank()
  )
```

## 6. Overall Relationship between car brands and claims

```{r}
# Aggregate data by categorical segments
treemap_data <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  group_by(vehicle_brand, municipality_type, circulation_area) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    # Calculate the Claim Rate for color-mapping
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  # Filter out tiny segments (e.g., fewer than 5 policies) to keep the visual clean
  filter(total_policies >= 5) %>%
  # Create a clean text label for each tile
  mutate(tile_label = paste0(vehicle_brand, "\n", municipality_type, "-", circulation_area))

ggplot(treemap_data, aes(area = total_policies, fill = claim_rate, label = tile_label)) +
  geom_treemap(color = "white", start = "topleft") +
  # Using the same intense color scale to easily contrast with your previous heatmap
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate",
                       labels = scales::percent_format(accuracy = 1)) +
  # Dynamic text wrapping inside the tiles
  geom_treemap_text(fontface = "bold", color = "white", place = "center", 
                    grow = FALSE, size = 11, min.size = 6) +
  labs(
    title = "Multivariate Portfolio Risk Treemap",
    subtitle = "Tile Size = Volume of Policies | Tile Color = Liability Property Claim Rate",
    caption = "Filtered for segments with 5+ policies"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```

Our high-volume segments (left) exhibit a stable, controlled claim rate of approximately 10-12%,  However, the treemap reveals significant risk volatility in our low-volume, niche segments (bottom-right), where claim rates swing drastically up to 25%. This indicates a need to either consolidate these micro-segments or apply stricter pricing loading to these volatile pockets."

## 7. Profile of Drivers for low profitability plans

In the previous section we determine that the above mentioned products warrants further study We will look into the profile of the drivers within this segments

### Age of low profitability plans

"We are testing whether our Comprehensive without Excess/Third-Party with two or more Cov/Basic Third Party portfolio is backed by the same age profile as our general business baseline ($H_0$), or if it has drifted into a statistically distinct demographic risk pool ($H_1$)." We will also use the median of the whole dataset driver's age as the test value.

::: panel-tabset
### Comprehensive without Excess

```{r}
ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Comprehensive without Excess") ,
  x = driver_age,
  type = "bayesian",      
  test.value = median(df_transformed$driver_age, na.rm = TRUE),  
  binwidth = 1,               
  title = glue("Statistical Evaluation of Driver Age Against Portfolio Median: {median(df_transformed$driver_age, na.rm = TRUE)}"),
  xlab = "Driver Age (Years)",
  ylab = "Count"
)
```

### Third-Party with two or more Cov

```{r}

ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Third-Party with two or more Cov") ,
  x = driver_age,
  type = "bayesian",      
  test.value = median(df_transformed$driver_age, na.rm = TRUE),  
  binwidth = 1,               
  title = glue("Statistical Evaluation of Driver Age Against Portfolio Median: {median(df_transformed$driver_age, na.rm = TRUE)}"),
  xlab = "Driver Age (Years)",
  ylab = "Count"
)
```

### Basic Third-Party

```{r}
ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Basic Third-Party") ,
  x = driver_age,
  type = "bayesian",      
  test.value = median(df_transformed$driver_age, na.rm = TRUE), 
  binwidth = 1,              
  title = glue("Statistical Evaluation of Driver Age Against Portfolio Median: {median(df_transformed$driver_age, na.rm = TRUE)}"),
  xlab = "Driver Age (Years)",
  ylab = "Count"
)
```
:::

From the above charts, we can conclude that with high statistical certainty that this sub-portfolio of plans has drifted into a structurally different demographic risk profile than the overall portfolio"

### Age of cars

::: panel-tabset
### Comprehensive without Excess

```{r}
ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Comprehensive without Excess") ,
  x = vehicle_age,
  type = "bayesian",      
  test.value = median(df_transformed$vehicle_age, na.rm = TRUE),  
  binwidth = 1,                    
  title = glue("Statistical Evaluation of Vehicle Age Against Portfolio Median: {median(df_transformed$vehicle_age, na.rm = TRUE)}"),
  xlab = "Vehicle Age (Years)",
  ylab = "Count"
)
```

### Third-Party with two or more Cov

```{r}
ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Third-Party with two or more Cov") ,
  x = vehicle_age,
  type = "bayesian",      
  test.value = median(df_transformed$vehicle_age, na.rm = TRUE),  
  binwidth = 1,                    
  title = glue("Statistical Evaluation of Vehicle Age Against Portfolio Median: {median(df_transformed$vehicle_age, na.rm = TRUE)}"),
  xlab = "Vehicle Age (Years)",
  ylab = "Count"
)
```

### Basic Third-Party

```{r}
ggstatsplot::gghistostats(
  data = df_transformed %>% filter(policy_type == "Basic Third-Party") ,
  x = vehicle_age,
  type = "bayesian",      
  test.value = median(df_transformed$vehicle_age, na.rm = TRUE),  
  binwidth = 1,                    
  title = glue("Statistical Evaluation of Vehicle Age Against Portfolio Median: {median(df_transformed$vehicle_age, na.rm = TRUE)}"),
  xlab = "Vehicle Age (Years)",
  ylab = "Count"
)
```
:::

Observation

Comprehensive without Exces

- While a baseline evaluation shows a vehicle age median close to the general portfolio target of 25 ($\hat{\mu}_{\text{MAP}} = 24.63$), Bayesian test metrics confirm a severe structural drift ($\log_{e}(\text{BF}_{01}) = -100.82$). The underlying data exhibits a highly pronounced bimodal distribution, splitting into a modern cohort (peaking at ~22 years) and a legacy asset cohort (peaking at ~36 years).

Third-Party with two or more Cov

- Bayesian test metrics confirm a severe and foundational demographic shift away from the portfolio median of 25 ($\log_{e}(\text{BF}_{01}) = -631.14$). The segment exhibits a highly pronounced bimodal distribution that heavily skews younger, with its primary risk peak established at 19 years ($\hat{\mu}_{\text{MAP}} = 19.02$). This massive structural drift indicates that the product line is predominantly absorbing a younger asset class, separated by a sharp structural valley from a secondary legacy fleet segment near 35 years. Standardized baseline underwriting assumptions must be updated for this product line to reflect this younger, highly distinct distribution profile

Basic Third-Party

- Bayesian analysis decisively rejects the assumption that this segment matches our baseline portfolio median of 25 ($\log_{e}(\text{BF}_{01}) = -75.69$), despite the empirical median matching the target almost exactly ($\hat{\mu}_{\text{MAP}} = 24.85$). The underlying distribution has fractured into a complex, multimodal profile consisting of an emerging entry-level fleet (~10 years), a dominant modern cohort peaking at ~20 years, and a highly volatile, fragmented legacy asset cohort spanning 34–45 years. Relying on aggregate point estimates (like the median) for this specific plan creates severe underwriting exposure, as it treats three distinct, highly polarized risk demographics as a single homogeneous group.

## Identifying the type of claims triggered

In this section, we want to understand for 'Third-Party with two or more Cov'/'Basic Third Party' portfolio. What are the claims most triggered?

::: panel-tabset
### Comprehensive without Excess

```{r}
df_claims_raw <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)


set.seed(123) 
df_sample <- df_claims_raw %>% 
  mutate(policy_row_index = row_number()) # Create a clean Y-axis identifier

#
df_heatmap_ready <- df_sample %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%

  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


ggplot(df_heatmap_ready, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +

  geom_tile() + 

  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +
  
  theme_minimal(base_size = 12) +
  labs(
    title = "Comprehensive without Excess Claims Activation Heatmap",
    subtitle = glue("Visual matrix audit of {nrow(df_sample)} claim-active policies"),
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )

```

### Third-Party with two or more Cov

```{r}

# 1. Isolate the subset and select the binary claim columns
df_claims_raw <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)


set.seed(123) 
df_sample <- df_claims_raw %>% 
  mutate(policy_row_index = row_number()) # Create a clean Y-axis identifier

#
df_heatmap_ready <- df_sample %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%

  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


ggplot(df_heatmap_ready, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +

  geom_tile() + 

  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +
  
  theme_minimal(base_size = 12) +
  labs(
    title = "Third-Party with two or more Coverage Claims Activation Heatmap",
    subtitle = glue("Visual matrix audit of {nrow(df_sample)} claim-active policies"),
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )



```

### Basic Third-Party

```{r}

df_claims_raw <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)


set.seed(123) 
df_sample <- df_claims_raw %>% 
  mutate(policy_row_index = row_number()) # Create a clean Y-axis identifier


df_heatmap_ready <- df_sample %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%

  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


ggplot(df_heatmap_ready, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +

  geom_tile() + 

  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +
  
  theme_minimal(base_size = 12) +
  labs(
    title = "Basic Third-Party Claims Activation Heatmap",
    subtitle = glue("Visual matrix audit of {nrow(df_sample)} claim-active policies"),
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )

```
:::

Comprehensive without Excess plan
- there is a prominent, severe vertical band of activation for Property Claims, followed closely by high activation in Glass and Liability Property claims.

Third-Party with Two or More Coverages
- there is a distinct, razor-sharp vertical concentration exclusively under Liability Property Claims and Liability Injury Claims. Other features like theft, fire, or glass show almost zero activation.

Basic Third party Claims
- Most claims are concentrated in liability protection. In spain. As expected as Basic Third party also known as [civil obligatoria](https://www.giambronelaw.com/site/advice/personal-injury-advice/personal-injury-spain/driving-in-spain/) covers legal protection, liability property and liability injury to the people or property caused by you.

Overview
liability property claims are the most common due to Spains' [compensation agreement](https://www.ofesauto.es/en/know-the-spanish-direct-compensation-agreement-cide-to-determine-which-vehicle-is-the-responsible/). This agreement makes liability property damage easy to file.

## 8. Correlation

::: panel-tabset
### Continous Variable

```{r}

 corr_data <- df_transformed %>%
  select(
    driver_age, 
    vehicle_age, 
    age_driving_licence,
    vehicle_value,
    total_premium, 
    power_to_weight_ratio,
    liability_property_claims  # Your target claim count metric
  ) %>%
  # Ensure all selected features are strictly evaluated as numeric matrices
  mutate(across(everything(), as.numeric)) %>%
  # Handle missing records safely to maintain clean row-counts
  mutate(across(everything(), ~tidyr::replace_na(.x, 0)))

# 2. Generate the statistical correlation matrix
ggcorrmat(
  data = corr_data,
  type = "spearman",          # Use "nonparametric" for Spearman or "bayesian" if preferred
  colors = c("#2c7bb6", "#ffffbf", "#d7191c"), # Classic blue-to-red diverging scale
  title = "Portfolio Risk Profile: Correlation Matrix Matrix",
  subtitle = "Evaluating linear relationships across driver metrics, vehicle assets, and claim outcomes",
  show.legend = TRUE,
  matrix.type = "upper",        # Shows top triangle to remove visual redundancy
  lab.size = 3.5,               # Size of correlation coefficients inside the squares
  ggtheme = ggplot2::theme_minimal()
)
```

### Categorical Variables

```{r}

# 1. Isolate and clean the factors
cramer_df <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  select(
    policy_status, business_type, payment_frequency, 
    bonus_score, fuel_type, vehicle_brand, 
    municipality_type, circulation_area
  ) %>%
  mutate(across(everything(), as.factor))

# 2. Compute the exact 8x8 Cramér's V matrix mathematically
cat_vars <- colnames(cramer_df)
n_vars <- length(cat_vars)
v_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars, dimnames = list(cat_vars, cat_vars))

for (i in 1:n_vars) {
  for (j in 1:n_vars) {
    # Generate cross-tabulation table
    contingency_table <- table(cramer_df[[cat_vars[i]]], cramer_df[[cat_vars[j]]])
    # Store the scalar Cramer's V value
    v_matrix[i, j] <- assocstats(contingency_table)$cramer
  }
}

# 3. Convert the matrix to a long data frame format for ggplot
v_long <- as.data.frame(v_matrix) %>%
  mutate(Variable1 = rownames(.)) %>%
  pivot_longer(cols = -Variable1, names_to = "Variable2", values_to = "CramersV")

# 4. Plot a clean, beautifully spaced association tile grid
ggplot(v_long, aes(x = Variable1, y = Variable2, fill = CramersV)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient2(low = "#ffffbf", mid = "#fdae61", high = "#d7191c", 
                       midpoint = 0.5, limit = c(0, 1), name = "Cramér's V") +
  geom_text(aes(label = round(CramersV, 2)), color = "black", fontface = "bold", size = 4) +
  labs(
    title = "Clean Categorical Feature Association Matrix",
    subtitle = "Calculated via global chi-square contingencies | Exactly 8x8 Dimensions",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```
:::


## 9.Profile of Age and Vehicle Age of selected Policy Type Claims

This section performs multi-variate analysis to understand the relationship between drivers age, vehicle age and claims.

::: panel-tabset
### Comprehensive without Excess

```{r}
heatmap_data <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  # Create logical bins for your continuous variables
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

# 2. Plot the binned matrix
ggplot(heatmap_data, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3.5, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Identifying localized risk spikes across driver and vehicle age pockets",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 11),
    panel.grid = element_blank()
  )
```

### Third-Party with two or more Cov

```{r}
heatmap_data <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  # Create logical bins for your continuous variables
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

# 2. Plot the binned matrix
ggplot(heatmap_data, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3.5, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Identifying localized risk spikes across driver and vehicle age pockets",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 11),
    panel.grid = element_blank()
  )
```

Our liability property claim spike is being driven by two specific operational pockets: middle-aged to senior drivers operating brand-new or modern vehicles (0–6 years old), alongside a high, persistent baseline risk from our under-35 cohort across all car ages.

### Basic Third-Party

```{r}

heatmap_data <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  # Create logical bins for your continuous variables
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

# 2. Plot the binned matrix
ggplot(heatmap_data, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3.5, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Identifying localized risk spikes across driver and vehicle age pockets",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 11),
    panel.grid = element_blank()
  )
```
:::



## 10. Premium Exposure By Drivers Age and Vehicle Age

In this section we will take into account the exposure column. Create variables to address the "Average Premium Revenue per Unit of Exposure Year" in section 1 and we will compare this new variable against Drivers Age and Vehicle Age.

This is meant to be use in conjunction with the heat map above

::: panel-tabset
### Comprehensive without Excess

```{r}
#| fig-width: 12
#| fig-height: 6
#| dpi: 150

## driver age
premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p1 <- ggplot(premium_exposure_analysis, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )

#vehicle age

premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p2 <- ggplot(premium_exposure_analysis, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )

p1 +p2
```

### Third-Party with two or more Cov

```{r}
#| fig-width: 12
#| fig-height: 6
#| dpi: 150

## driver age
premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p1 <- ggplot(premium_exposure_analysis, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )

#vehicle age

premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p2 <- ggplot(premium_exposure_analysis, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d62728")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )

p1 +p2
```

### Basic Third Party

```{r}
#| fig-width: 12
#| fig-height: 6
#| dpi: 150
# Driver age
premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p1 <- ggplot(premium_exposure_analysis, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )

## Vehicle age

premium_exposure_analysis <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

# 1. Rename the variable to match its true mathematical scope
plan_avg_premium <- sum(premium_exposure_analysis$total_premium) / sum(premium_exposure_analysis$total_exposure)

# 2. Update the conditional logic flag
premium_exposure_analysis <- premium_exposure_analysis %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

# 3. Update the ggplot layers to use the correct labels
p2 <-ggplot(premium_exposure_analysis, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  
  # Update the threshold line target
  geom_hline(yintercept = plan_avg_premium, linetype = "dashed", color = "gray40", linewidth = 1) +
  
  # Update the text label mapping onto the chart space
  annotate("text", x = 1.5, y = plan_avg_premium + 15, 
           label = paste0("    Plan Average Baseline: $", round(plan_avg_premium, 2)), 
           color = "gray30", fontface = "italic", size = 3.5) +
  
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Average Premium Revenue per Unit of Exposure Year",
    subtitle = "Slicing Third-Party with two or more Cov plan to spot sub-group underpricing",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  )


p1 + p2
```
:::

Comprehensive without Excess:
There is a distinct pricing mismatch within the Third-Party with two or more Coverages portfolio. While underwriting successfully charges a steep risk premium on young drivers (18–25) and brand-new vehicles (0–3 years), it severely underprices the mid-aged driver segment (36–55) operating mid-age vehicles (4–6 years).

This exact underpriced demographic is likely to be driving the portfolio's exposure, peaking at a 37.1% claim frequency. To stabilize the portfolio's margin, the underwriting team should review the pricing tiers upward for drivers aged 36–55 who operate vehicles under 6 years old.


Third-Party with two or more coverage:
While the pricing structure successfully accounts for high-risk young drivers (18–25), it completely fails to protect against a major risk spike in senior operators.

Drivers aged 66+ operating 4–6 year old vehicles generate the portfolio's highest liability claim rate at 26.1%, yet this demographic pays below-average baseline premiums. Because basic plans have thin margins by design, this unpriced exposure severely compromises the product's profitability.

Basic Third Party:
The underwriter successfully priced in premiums for the 18-25 age groups with newer vehicles from 0 -6 years.

However, a critical underwriting leakage exists in the mature segment: drivers aged 56–65 operating 7–10 year old cars run a highly elevated 21.4% claim rate, yet they are granted below-average premium rates.
