---
title: "Take Home Exercise 1 Executive Summary"
author: "Alvin Ong Wei Ming"
date: "May 15, 2026"
date-modified: "last-modified"
format: 
  revealjs:
    logo: motor_insurance_logo.png
    theme: simple        
    code-fold: true      
    echo: false
execute:
  eval: true            
  warning: false        
  freeze: true          
---

## Content page{.smaller}

-   1.Financial health of the business
-   2.Product profitability
-   3.Total Premiums payment over the years
-   4.Total Premiums payment over the years \n Third-Party with Glass Cov
-   5.Total Premiums payment over the years \n Basic Third-Party
-   6.Types of claims per low profitability portfolios
-   7.Profile of Driver and Vehicle age against claims: Comprehensive without Excess 
-   8.Profile of Driver and Vehicle age against claims: Third Party with two or more coverage 
-   9.Profile of Driver and Vehicle age against claims: Basic Third Party coverage
-   10.Vehicle Type and District claim rate
-   11.Data Issues
-   12.Conclusion

```{r}
pacman::p_load(ggstatsplot, 
               ggdist,
               see,
               tidyverse,
               skimr,
               naniar,
               scales,
               arrow,
               patchwork,
               GGally,
               glue,
               ggpubr,
               treemapify,
               vcd)

df_transformed <- read_parquet("df_transformed.parquet")


plot_policy_distribution <- function(data, target_policy, target_metric = "total_premium") {
  
  # 1. Dynamically filter and prepare data based on function arguments
  plot_data <- data %>%
    filter(
      policy_type == target_policy,
      .data[[target_metric]] > 0
    ) %>%
    mutate(year = factor(year))
  
  # Determine dynamic text for labels
  metric_label <- gsub("_", " ", target_metric) %>% tools::toTitleCase()
  
  # 2. Safe Zero-Data Check (Replacing the Kruskal check)
  if (nrow(plot_data) == 0) {
    warning(paste0("No data found for policy type: '", target_policy, "' with ", metric_label, " > 0."))
    return(NULL)
  }
  
  # 3. Generate the Plot
  p <- ggplot(plot_data, aes(x = year, y = .data[[target_metric]], fill = year)) +
    
    # The Half-Eye distribution view
    stat_halfeye(
      adjust = 0.5,
      width = 0.6,
      .width = c(0.66, 0.95),
      point_color = "black",
      interval_color = "black",
      alpha = 0.7
    ) +
    
    # Scale and formatting
    scale_y_log10(labels = scales::label_comma(prefix = "$")) +
    scale_fill_brewer(palette = "Blues", guide = "none") +
    
    # Informative labels without the statistical test string
    labs(
      title = paste0(metric_label, " Distribution across Years"),
      subtitle = target_policy,
      x = "Year",
      y = paste0(metric_label, " (Log Scale)")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(face = "bold")
    ) +
    coord_flip() 
  
  return(p)
}


plot_claims_distribution <- function(data, target_policy, target_metric = "total_incurred") {
  
  # 1. Dynamically filter and prepare data based on function arguments
  # We use .data[[target_metric]] to allow passing the column name as a string
  plot_data <- data %>%
    filter(
      policy_type == target_policy,
      .data[[target_metric]] > 0
    ) %>%
    mutate(year = factor(year))
  
  # Determine dynamic text for labels
  metric_label <- gsub("_", " ", target_metric) %>% tools::toTitleCase()
  
  # 2. Safe Zero-Data Check (Replacing the Kruskal check)
  if (nrow(plot_data) == 0) {
    warning(paste0("No data found for policy type: '", target_policy, "' with ", metric_label, " > 0."))
    return(NULL)
  }
  
  # 3. Generate the Plot
  p <- ggplot(plot_data, aes(x = year, y = .data[[target_metric]], fill = year)) +
    
    # The Half-Eye distribution view
    stat_halfeye(
      width = 0.6,
      .width = c(0.66, 0.95),
      point_color = "black",
      interval_color = "black",
      alpha = 0.7
    )  + 
    
    # Scale and formatting
    scale_y_log10(
      limits = c(0.10, 500000), 
      labels = scales::label_dollar(),
      breaks = c(0.10, 1, 10, 100, 1000, 10000, 100000, 500000)
    ) +
    scale_fill_brewer(palette = "Reds", guide = "none") +
    
    # Informative labels that update dynamically based on your inputs
    labs(
      title = paste0(metric_label, " Distribution across Years"),
      subtitle = target_policy,
      x = "Year",
      y = paste0(metric_label, " (Log Scale)")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(face = "bold")
    ) +
    coord_flip() 
  
  return(p)
}

```

::: notes

:::

## **1. Financial health of the business**

```{r}
#| echo: false
#| fig-width: 16
#| fig-height: 8
yearly_summary <- df_transformed %>%
  group_by(year) %>%
  summarize(
    sum_premium = sum(total_premium, na.rm = TRUE),
    sum_incurred = sum(total_incurred, na.rm = TRUE),
    underwriting_profit = sum_premium - sum_incurred,
    avg_rev_per_exposure = sum_premium / sum(total_exposure, na.rm = TRUE)
  ) %>%
  arrange(year) %>%
  # Calculate YoY changes
  mutate(
    premium_yoy = (sum_premium - lag(sum_premium)) / lag(sum_premium) * 100,
    incurred_yoy = (sum_incurred - lag(sum_incurred)) / lag(sum_incurred) * 100
  )

# =========================================================================
# STEP 2: Pivot & Format Labels for Plot 1 (Grouped Bars)
# =========================================================================
p1_data <- yearly_summary %>%
  select(year, `Total Premium` = sum_premium, `Total Incurred Claims` = sum_incurred, premium_yoy, incurred_yoy) %>%
  pivot_longer(
    cols = c(`Total Premium`, `Total Incurred Claims`), 
    names_to = "Metric", 
    values_to = "Amount"
  ) %>%
  # Assign the correct YoY percentage to the correct metric row
  mutate(
    yoy_val = if_else(Metric == "Total Premium", premium_yoy, incurred_yoy),
    # Format text: If it's the first year, just show amount. Otherwise, show amount + YoY %
    label_text = case_when(
      is.na(yoy_val) ~ label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount),
      yoy_val >= 0   ~ paste0(label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount), "\n(+", round(yoy_val, 1), "%)"),
      TRUE           ~ paste0(label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(Amount), "\n(", round(yoy_val, 1), "%)")
    )
  )

# =========================================================================
# STEP 3: Build Plot 1 - Grouped Bar Chart with Value & % Labels
# =========================================================================
# Explicitly defining bar positioning width so the text alignments align flawlessly
bar_width <- 0.6
dodge_width <- 0.7

p1 <- ggplot(p1_data, aes(x = factor(year), y = Amount, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(width = dodge_width), width = bar_width) +
  
  # Add dynamic text labels on top of bars
  geom_text(
    aes(label = label_text),
    position = position_dodge(width = dodge_width),
    vjust = -0.25,       
    size = 3,            
    fontface = "bold",
    lineheight = 0.85    
  ) +
  
  scale_fill_manual(values = c("Total Premium" = "#2ca02c", "Total Incurred Claims" = "#d9534f")) +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0, 0.15)) # Adds 15% headroom at the top so labels don't clip
  ) +
  labs(
    title = "A: Premium Inflow vs. Claims Outflow",
    x = NULL, 
    y = "Total Financial Amount (Y-O-Y % Increase)", 
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 4: Build Plot 2 - Net Underwriting Profit Trend with Labels
# =========================================================================
yearly_summary <- yearly_summary %>%
  mutate(
    profit_label = label_number(accuracy = 0.1, scale_cut = cut_short_scale(), prefix = "$")(underwriting_profit)
  )

p2 <- ggplot(yearly_summary, aes(x = factor(year), y = underwriting_profit, fill = underwriting_profit > 0)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.9) +
  geom_text(
    aes(label = profit_label, vjust = if_else(underwriting_profit >= 0, -0.4, 1.4)),
    size = 3,
    fontface = "bold"
  ) +
  scale_fill_manual(values = c("TRUE" = "#2ca02c", "FALSE" = "#d9534f"), guide = "none") +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0.15, 0.15))
  ) +
  labs(
    title = "B: Net Underwriting Profit Trend",
    subtitle = "Green = Profitable | Red = Underwriting Loss",
    x = NULL, 
    y = "Net Profit"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 9, color = "gray30"),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 5: Build Plot 3 - Average Revenue Per Unit of Exposure
# =========================================================================
p3 <- ggplot(yearly_summary, aes(x = factor(year), y = avg_rev_per_exposure, group = 1)) +
  geom_line(color = "#D4AF37", linewidth = 1.2) +
  geom_point(color = "#1c3b57", size = 3) +
  geom_text(
    aes(label = paste0("$", round(avg_rev_per_exposure, 1))), 
    vjust = -1, 
    size = 3.5, 
    fontface = "bold"
  ) +
  scale_y_continuous(
    labels = label_number(prefix = "$"),
    expand = expansion(mult = c(0.1, 0.2))
  ) +
  labs(
    title = "C: Average Premium Revenue per Unit of Exposure Year",
    x = "Year", 
    y = "Avg Premium / Exposure"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank()
  )

# =========================================================================
# STEP 6: Assemble Framework Dashboard Layout and Print
# =========================================================================
dashboard <- (p1 | p2) / p3 + 
  plot_annotation(
    title = "Overall Financial Health",
    subtitle = "Year-on-Year metrics tracking revenue growth, claims impact, and net margin",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, margin = margin(b = 5)),
      plot.subtitle = element_text(color = "gray25", size = 11, margin = margin(b = 15))
    )
  )

# Preview the clean dashboard
print(dashboard)
```
::: notes
- On the top left we can observed the premiums and claims outflow from the business.

- On the top right, we also observe the net profit margin has slowed significantly since 2023 to 2024

- Lastly, we are receiving lower average premiums per exposure year. This signals we need to relook at how we are pricing our policies
:::

## **2. Product profitability**

```{r}
premium_summary <- df_transformed %>%
  group_by(policy_type) %>%
  summarize(total_premium_sum = sum(total_premium, na.rm = TRUE))


incurred_summary <- df_transformed %>%
  group_by(policy_type) %>%
  summarize(total_incurred_sum = sum(total_incurred, na.rm = TRUE))

# Merge and calculate cumulative metrics based on Premium Rank
pareto_combined <- premium_summary %>%
  left_join(incurred_summary, by = "policy_type") %>%
  arrange(desc(total_premium_sum)) %>%
  mutate(
    policy_type = factor(policy_type, levels = policy_type),
    
    # Premium Tail Metrics
    cum_sum_prem = cumsum(total_premium_sum),
    total_sum_prem = sum(total_premium_sum),
    cum_pct_prem = (cum_sum_prem / total_sum_prem) * 100,
    remaining_pct_prem = 100 - lag(cum_pct_prem, default = 0),
    
    # Claims Tail Metrics
    cum_sum_inc = cumsum(total_incurred_sum),
    total_sum_inc = sum(total_incurred_sum),
    cum_pct_inc = (cum_sum_inc / total_sum_inc) * 100,
    remaining_pct_inc = 100 - lag(cum_pct_inc, default = 0)
  )

# Master anchors for dual-axis transformation
total_premium_total <- sum(pareto_combined$total_premium_sum)
max_y_limit <- total_premium_total * 1.1

# Pivot long for side-by-side dodge bars
bar_data <- pareto_combined %>%
  select(policy_type, Premium = total_premium_sum, Claims = total_incurred_sum) %>%
  pivot_longer(cols = c(Premium, Claims), names_to = "Metric", values_to = "Amount") %>%
  mutate(Metric = factor(Metric, levels = c("Premium", "Claims")))

# --- 2. GENERATE OVERLAYED PLOT ---
ggplot() +
  # Side-by-side grouped bars for Premium vs Claims
  geom_bar(data = bar_data, aes(x = policy_type, y = Amount, fill = Metric), 
           stat = "identity", position = position_dodge(width = 0.8), alpha = 0.85, width = 0.7) +
  
  # Gold Line: Remaining Revenue Tail
  geom_line(data = pareto_combined, aes(x = as.numeric(policy_type) - 0.2, 
                                        y = remaining_pct_prem * (total_premium_total / 100), group = 1, color = "Remaining Revenue Tail"), 
            linewidth = 1.2) +
  geom_point(data = pareto_combined, aes(x = as.numeric(policy_type) - 0.2, 
                                         y = remaining_pct_prem * (total_premium_total / 100)), color = "#D4AF37", size = 2.5) +
  
  # Dual Axis Configuration
  scale_y_continuous(
    name = "Financial Amount ($)",
    limits = c(0, max_y_limit),
    breaks = seq(0, max_y_limit, by = 10000000),
    labels = label_comma(),
    sec.axis = sec_axis(
      trans = ~ . / total_premium_total * 100,
      name = "Remaining Portfolio Percentage (%)",
      breaks = seq(0, 100, by = 10)
    )
  ) +
  
  # Styling and Legends
  scale_fill_manual(values = c("Premium" = "#2ca02c", "Claims" = "#d9534f")) +
  scale_color_manual(values = c("Remaining Revenue Tail" = "#D4AF37", "Remaining Claims Tail" = "grey25")) +
  
  labs(
    title = "Portfolio Premium & Incurred Claims",
    subtitle = "Bars show Total Premium (Green)/ Claims (Red).\n Gold Line shows the propotion of portfolio.",
    x = "Policy Type",
    fill = "Financial Metrics",
    color = "Portfolio Tails"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  )
```
::: notes
- In this slide, we observed 2 critical observation.

A performance review of the product portfolio reveals a dangerous concentration risk, with over 80% of total premiums heavily reliant on just two offerings: Third Party (2 or more coverage) and Comprehensive with Excess.

The remaining product lines demonstrate severe profitability distress.

Comprehensive without Excess is currently operating at an breakeven. 
Concurrently, Third Party with Glass Coverage and basic Third Party plans are yielding razor-thin margins, failing to contribute meaningfully to the bottom line despite their operational footprint. 

We will further investigate the cause for these 3 portfolios in subsequent slides.
:::

## **3. Total Premiums payment over the years \n Comprehensive without Excess**



```{r}
#| fig-width: 16
#| fig-height: 8

p1 <- plot_policy_distribution(df_transformed, "Comprehensive without Excess")
p2 <- plot_claims_distribution(df_transformed,"Comprehensive without Excess")

p1 + p2
```


::: notes

Note: The scale is in log, but the x axis is in actual dollars

Key insight:
Comprehensive insurance without excess shows a stable distribution for both premiums and claims from 2022 through 2024, explaining why it is barely making a profit on an aggregate level. Crucially, the multi-peaked distributions across all years confirm the existence of distinct customer sub-groups (e.g., a distinct low-premium tier and a high-premium tier).

Because year-over-year premium growth has remained stagnant, strategic adjustments are required. Further monitoring and tweaking of this policy are needed to intentionally pull more customers into the higher-tier, higher-margin brackets to improve profitability."
:::

## **4. Total Premiums payment over the years \n Third-Party with Glass Cov**

```{r}
#| fig-width: 16
#| fig-height: 8
p1 <- ggplot(df_transformed %>% filter(policy_type == "Third-Party with Glass Cov"), 
       aes(x = factor(year), y = total_premium, fill = factor(year))) +
  
  # Keeps your original smooth halfeye
  ggdist::stat_halfeye(
    adjust = 0.5,
    width = 0.6,
    .width = c(0.66, 0.95),
    point_color = "black",
    interval_color = "black",
    alpha = 0.7
  ) +
  
  scale_y_continuous(labels = scales::label_comma(prefix = "$")) +

  coord_cartesian(ylim = c(0, 200)) + 
  
  scale_fill_brewer(palette = "Blues", guide = "none") +
  
  labs(
    title = "Total Premium Distribution across Years",
    subtitle = "Third-Party with Glass Cov",
    x = "Year",
    y = "Total Premium"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray30", size = 10, face = "italic"),
    panel.grid.minor = element_blank(),
    axis.text = element_text(face = "bold")
  ) +
  
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() 
p2 <- plot_claims_distribution(df_transformed,"Third-Party with Glass Cov")

 p1+p2
```
::: notes
Note: The scale is in log, but the x axis is in actual dollars

Third-Party with Glass Coverage shows an increasing volume of lower-priced premiums over the years, though the median premium remains stable around $200 On the claims side, while the median has held steady at approximately $1,000 from 2022 to 2024, the structure of the risk has shifted significantly. Claims have become much more heavily concentrated right at the $1,000 mark in 2024, alongside a wider variation of smaller claims emerging below the 25th percentile.

Based on the shifting claims structure, it is highly recommended to conduct a targeted risk and underwriting review focused on two areas
:::


## **5. Total Premiums payment over the years \n Basic Third-Party**

```{r}
#| fig-width: 16
#| fig-height: 8
p1 <- plot_policy_distribution(df_transformed, "Basic Third-Party")

p2 <- plot_claims_distribution(df_transformed,"Basic Third-Party")

p1+p2
```

::: notes
Note: The scale is in log, but the x axis is in actual dollars

The premium structure for this plan experienced significant compression in 2024, shifting away from the wider variations of previous years into a distinct bimodal distribution.

This structural change strongly indicates the emergence of two polarized risk segments or sub-groups within the same product line, even though the overall average premium level has stabilized between 2023 and 2024.

Concurrently, the portfolio's claim experience underwent a critical upward migration. While prior years were dominated by low-value "background" claims, 2024 losses heavily bottlenecked near a much higher median payout while maintaining highly volatile outer boundaries. 

This combination of a rising baseline cost per claim warrants a review of this insurance policy.
:::

## **6. Types of claims per low profitability portfolios**

```{r}
#| fig-width: 15
#| fig-height: 4.5
df_claims_p1 <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)

set.seed(123) 
df_sample_p1 <- df_claims_p1 %>% 
  mutate(policy_row_index = row_number())

df_heatmap_p1 <- df_sample_p1 %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%
  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


p1 <- ggplot(df_heatmap_p1, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +
  geom_tile() + 
  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +  
  theme_minimal(base_size = 12) +
  labs(
    title = "Comprehensive without Excess Claims",
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )



df_claims_p2 <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)

set.seed(123) 
df_sample_p2 <- df_claims_p2 %>% 
  mutate(policy_row_index = row_number())

df_heatmap_p2 <- df_sample_p2 %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%
  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


p2 <- ggplot(df_heatmap_p2, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +
  geom_tile() + 
  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +  
  theme_minimal(base_size = 12) +
  labs(
    title = "Third-Party with two or more Coverage Claims",
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )


df_claims_p3 <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  select(
    insured_id, year,
    liability_property_claims, liability_injury_claims,
    property_claims, theft_claims, fire_claims,
    glass_claims, legal_protection_claims, occupants_claims
  ) %>%
  filter(rowSums(select(., ends_with("_claims"))) > 0)

set.seed(123) 
df_sample_p3 <- df_claims_p3 %>% 
  mutate(policy_row_index = row_number())

df_heatmap_p3 <- df_sample_p3 %>%
  pivot_longer(
    cols = ends_with("_claims"), 
    names_to = "Claim_Type", 
    values_to = "Claim_Present"
  ) %>%
  mutate(
    Claim_Type = str_remove(Claim_Type, "_claims"),
    Claim_Type = str_replace_all(Claim_Type, "_", " "),
    Claim_Status = ifelse(Claim_Present == 1, "Claim Filed (1)", "No Claim (0)")
  )


p3 <- ggplot(df_heatmap_p3, aes(x = Claim_Type, y = factor(policy_row_index), fill = Claim_Status)) +
  geom_tile() + 
  scale_fill_manual(values = c("Claim Filed (1)" = "#E46726", "No Claim (0)" = "#F0F0F0")) +  
  theme_minimal(base_size = 12) +
  labs(
    title = "Basic Third-Party Claims",
    x = "Coverage Type",
    y = "",
    fill = "Status"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "white", linewidth = 0.5),
    legend.position = "top"
  )

p1 + p2 + p3
```
::: notes

The Comprehensive without Excess plan exhibits widespread, high-frequency claims, with severe activation bands across Property, Glass, and Liability Property claims

In stark contrast, Third-Party with Two or More Coverages shows a highly localized, sparse risk profile, with claims strictly concentrated within third-party property and bodily injury liabilities, indicating a lower-frequency but potentially higher-severity exposure.

For Basic Third-Party policies, claims are almost exclusively restricted to a vertical band of liability protection. This pattern aligns precisely with Spanish legal mandates, where entry-level insurance operates under Responsabilidad Civil Obligatoria
:::

## **7. Profile of Driver and Vehicle age against claims \nComprehensive without Excess **

```{r}
#| fig-width: 15
#| fig-height: 4.5
#| out-width: "100%"

# =========================================================================
# CHART 1: Liability Claim Frequency Matrix (Heatmap)
# =========================================================================
heatmap_data_comp <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

p1 <- ggplot(heatmap_data_comp, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Risk spikes across driver and vehicle age segments",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text = element_text(size = 9),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )

# =========================================================================
# CHART 2: Premium Revenue by Driver Age
# =========================================================================
premium_driver_comp <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_driver_comp <- sum(premium_driver_comp$total_premium) / sum(premium_driver_comp$total_exposure)

premium_driver_comp_final <- premium_driver_comp %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_driver_comp, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p2 <- ggplot(premium_driver_comp_final, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_driver_comp, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3.5, y = plan_avg_premium_driver_comp + 25, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_driver_comp, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Driver Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )

# =========================================================================
# CHART 3: Premium Revenue by Vehicle Age
# =========================================================================
premium_vehicle_comp <- df_transformed %>%
  filter(policy_type == "Comprehensive without Excess") %>%
  mutate(
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_vehicle_comp <- sum(premium_vehicle_comp$total_premium) / sum(premium_vehicle_comp$total_exposure)

premium_vehicle_comp_final <- premium_vehicle_comp %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_vehicle_comp, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p3 <- ggplot(premium_vehicle_comp_final, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_vehicle_comp, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3, y = plan_avg_premium_vehicle_comp + 25, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_vehicle_comp, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Vehicle Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )

# =========================================================================
# EXECUTE DOCK SIDE-BY-SIDE PANELS
# =========================================================================
p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```
:::notes
There is a distinct pricing mismatch within the Third-Party with two or more Coverages portfolio. While underwriting successfully charges a steep risk premium on young drivers (18–25) and brand-new vehicles (0–3 years), it severely underprices the mid-aged driver segment (36–55) operating mid-age vehicles (4–6 years).

This exact underpriced demographic is likely to be driving the portfolio's exposure, peaking at a 37.1% claim frequency. To stabilize the portfolio's margin, the underwriting team should review the pricing tiers upward for drivers aged 36–55 who operate vehicles under 6 years old.

:::

## **8. Profile of Driver and Vehicle age against claims \n Third Party with two or more coverage **


```{r}
#| fig-width: 15
#| fig-height: 4.5
#| out-width: "100%"


heatmap_data_tp2 <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

p1 <- ggplot(heatmap_data_tp2, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Risk spikes across driver and vehicle age segments",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text = element_text(size = 9),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )


premium_driver_tp2 <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_driver_tp2 <- sum(premium_driver_tp2$total_premium) / sum(premium_driver_tp2$total_exposure)

premium_driver_tp2_final <- premium_driver_tp2 %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_driver_tp2, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p2 <- ggplot(premium_driver_tp2_final, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_driver_tp2, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3.5, y = plan_avg_premium_driver_tp2 + 15, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_driver_tp2, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Driver Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )

# =========================================================================
# CHART 3: Premium Revenue by Vehicle Age
# =========================================================================
premium_vehicle_tp2 <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  mutate(
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_vehicle_tp2 <- sum(premium_vehicle_tp2$total_premium) / sum(premium_vehicle_tp2$total_exposure)

premium_vehicle_tp2_final <- premium_vehicle_tp2 %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_vehicle_tp2, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p3 <- ggplot(premium_vehicle_tp2_final, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_vehicle_tp2, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3, y = plan_avg_premium_vehicle_tp2 + 15, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_vehicle_tp2, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Vehicle Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )

# =========================================================================
# ASSEMBLE PANEL SIDE-BY-SIDE
# =========================================================================
p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
  
```

::: notes
While the pricing structure successfully accounts for high-risk young drivers (18–25), it completely fails to protect against a major risk spike in senior operators.

Drivers aged 66+ operating 4–6 year old vehicles generate the portfolio's highest liability claim rate at 26.1%, yet this demographic pays below-average baseline premiums. Because basic plans have thin margins by design, this unpriced exposure severely compromises the product's profitability.
:::



## **9. Profile of Driver and Vehicle age against claims \n Basic Third party **

```{r}
#| fig-width: 15
#| fig-height: 4.5
#| out-width: "100%"

heatmap_data_basic <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")),
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))
  ) %>%
  group_by(age_bin, v_age_bin) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  filter(total_policies > 10)

p1 <- ggplot(heatmap_data_basic, aes(x = v_age_bin, y = age_bin, fill = claim_rate)) +
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate") +
  geom_text(aes(label = scales::percent(claim_rate, accuracy = 0.1)), 
            color = "white", size = 3, fontface = "bold") +
  labs(
    title = "Liability Claim Frequency Matrix",
    subtitle = "Risk spikes across driver and vehicle age segments",
    x = "Vehicle Age Group",
    y = "Driver Age Group"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text = element_text(size = 9),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )


premium_driver_basic <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>%
  mutate(
    age_bin = cut(driver_age, 
                  breaks = c(17, 25, 35, 45, 55, 65, Inf), 
                  labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"))) %>%
  group_by(age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_driver_basic <- sum(premium_driver_basic$total_premium) / sum(premium_driver_basic$total_exposure)

premium_driver_basic_final <- premium_driver_basic %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_driver_basic, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p2 <- ggplot(premium_driver_basic_final, aes(x = age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_driver_basic, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3.5, y = plan_avg_premium_driver_basic + 15, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_driver_basic, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Driver Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Driver Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )


premium_vehicle_basic <- df_transformed %>%
  filter(policy_type == "Basic Third-Party") %>% 
  mutate(
    v_age_bin = cut(vehicle_age, 
                    breaks = c(-1, 3, 6, 10, 15, Inf), 
                    labels = c("0-3 yrs", "4-6 yrs", "7-10 yrs", "11-15 yrs", "16+ yrs"))) %>%
  group_by(v_age_bin) %>%
  summarise(
    total_premium  = sum(total_premium, na.rm = TRUE),
    total_exposure = sum(liability_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(total_exposure > 0) %>%
  mutate(avg_premium_per_exposure = total_premium / total_exposure)

plan_avg_premium_vehicle_basic <- sum(premium_vehicle_basic$total_premium) / sum(premium_vehicle_basic$total_exposure)

premium_vehicle_basic_final <- premium_vehicle_basic %>%
  mutate(
    pricing_status = ifelse(
      avg_premium_per_exposure >= plan_avg_premium_vehicle_basic, 
      "Above/At Plan Average", 
      "Below Plan Average"
    )
  )

p3 <- ggplot(premium_vehicle_basic_final, aes(x = v_age_bin, y = avg_premium_per_exposure, fill = pricing_status)) +
  geom_col(width = 0.6, color = "white", lwd = 0.5) +
  geom_hline(yintercept = plan_avg_premium_vehicle_basic, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 3, y = plan_avg_premium_vehicle_basic + 15, 
           label = paste0("Avg Baseline: $", round(plan_avg_premium_vehicle_basic, 2)), 
           color = "gray30", fontface = "italic", size = 2.8) +
  scale_fill_manual(
    values = c("Above/At Plan Average" = "#2ca02c", "Below Plan Average" = "#d9534f")
  ) +
  labs(
    title = "Premium Revenue by Vehicle Age",
    subtitle = "Spotting sub-group underpricing segments",
    x = "Vehicle Age Group",
    y = "Premium Revenue / Exposure Year ($)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    plot.subtitle = element_text(size = 8, color = "gray30"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "bottom"
  )

# =========================================================================
# ASSEMBLE PANEL SIDE-BY-SIDE
# =========================================================================
p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.position = "bottom")

```
::: notes

The underwriter successfully priced in premiums for the 18-25 age groups with newer vehicles from 0 -6 years.

However, a critical underwriting leakage exists in the mature segment: drivers aged 56–65 operating 7–10 year old cars run a highly elevated 21.4% claim rate, yet they are granted below-average premium rates.

:::


## 10. Vehicle Type and District claim rate

```{r}
# Aggregate data by categorical segments
treemap_data <- df_transformed %>%
  filter(policy_type == "Third-Party with two or more Cov") %>%
  group_by(vehicle_brand, municipality_type, circulation_area) %>%
  summarise(
    total_policies = n(),
    total_liability_claims = sum(liability_property_claims, na.rm = TRUE),
    # Calculate the Claim Rate for color-mapping
    claim_rate = total_liability_claims / total_policies,
    .groups = "drop"
  ) %>%
  # Filter out tiny segments (e.g., fewer than 5 policies) to keep the visual clean
  filter(total_policies >= 5) %>%
  # Create a clean text label for each tile
  mutate(tile_label = paste0(vehicle_brand, "\n", municipality_type, "-", circulation_area))

ggplot(treemap_data, aes(area = total_policies, fill = claim_rate, label = tile_label)) +
  geom_treemap(color = "white", start = "topleft") +
  # Using the same intense color scale to easily contrast with your previous heatmap
  scale_fill_viridis_c(option = "inferno", direction = -1, name = "Claim Rate",
                       labels = scales::percent_format(accuracy = 1)) +
  # Dynamic text wrapping inside the tiles
  geom_treemap_text(fontface = "bold", color = "white", place = "center", 
                    grow = FALSE, size = 11, min.size = 6) +
  labs(
    title = "Multivariate Portfolio Risk Treemap",
    subtitle = "Tile Size = Volume of Policies | Tile Color = Liability Property Claim Rate",
    caption = "Filtered for segments with 5+ policies"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```

:::notes
Our high-volume segments (left) exhibit a stable, controlled claim rate of approximately 10-12%,  However, the treemap reveals significant risk volatility in our low-volume, niche segments (bottom-right), where claim rates swing drastically up to 25%. This indicates a need to either consolidate these micro-segments or apply stricter pricing loading to these volatile pockets."
:::

## **11. Data Issues**

```{r}
df <-read.csv2("Dataset of motor insurance portfolio.csv")
  
gg_miss_upset(df, nsets = 4)
```
:::notes
Missing values in vehicle_value and fuel_type stem from systematic pipeline anomalies rather than random omissions. Engineering investigation is underway; missing fields were temporarily imputed to preserve data integrity for modeling
:::

## 12. Conclusion{.smaller .scrollable}

### Overall Findings 

- Over 80% of total premium revenue is reliant on just two offerings: Third-Party two or more coverages and Comprehensive with Excess.

- Year-on-year premium growth has flattened, and we are receiving a lower average premium per exposure year across the business, signaling an urgent need to relook at baseline pricing.

### Plug Critical Underwriting Price Leakages

Granular subgroup analysis reveals that while high-risk young drivers ($18\text{--}25$) are priced correctly, we are severely underpricing specific high-risk mature demographics:

- Drivers aged $36\text{--}55$ operating vehicles aged $4\text{--}6$ years peak at a massive 37.1% claim frequency but are currently charged below-average premiums.

- Senior operators aged $66+$ driving vehicles aged $4\text{--}6$ years generate a 26.1% claim rate while paying low baseline premiums.The Mature Leakage: Drivers aged $56\text{--}65$ with vehicles aged $7\text{--}10$ years run an elevated 21.4% claim rate but are granted below-average premium rates.

- Drivers aged $56\text{--}65$ with vehicles aged $7\text{--}10$ years run an elevated 21.4% claim rate but are granted below-average premium rates

### Data Issues

- To partnering with the engineering team to resolve these collection and validation errors pertaining to 1,287 missing fuel_type records and 513 missing vehicle_value records, ensuring data integrity for future analysis
