---
title: "EDA"
format: html
---

```{r}
pacman::p_load(plotly, crosstalk, DT, 
               ggdist, ggridges, colorspace,
               gganimate, tidyverse,skimr,naniar,glimpse)

```


## Read in Data
```{r}
df <-read.csv2("C:/BubblyCoder/ISSS608-VAA/Take_Home_Ex/Dataset of motor insurance portfolio.csv")
head(df)
```


## See overall and missing values.
```{r}

#notice that fuel type has 1287 missing values
# vehicle type has 513 value
# vehicle_age has 2 missing value
# age_driving_license has 2 missing values
skim(df)
```
## see Data types'
```{r}
str(df)

```

## data type conversion
We also created an indicator called underwriting profit to identify which portfolios are losses.
```{r}
df <- df %>% mutate(
    # all the factor like variables
    policy_type = as.factor(policy_type),
    policy_status = as.factor(policy_status),
    business_type = as.factor(business_type),
    payment_frequency = as.factor(bonus_score),
    fuel_type = as.factor(fuel_type),
    municipality_type = as.factor(municipality_type),
    circulation_area = as.factor(circulation_area),
    
    # these are all the integer like variables
    vehicle_value = as.numeric(vehicle_value),
    power_to_weight_ratio = as.numeric(power_to_weight_ratio),
    total_premium = as.numeric(total_premium),
    liability_premium = as.numeric(liability_premium),
    property_damage_premium = as.numeric(property_damage_premium),
    theft_premium = as.numeric(theft_premium),
    fire_premium = as.numeric(fire_premium),
    glass_premium = as.numeric(glass_premium),
    legal_protection_premium = as.numeric(legal_protection_premium),
    occupants_premium = as.numeric(occupants_premium),
    total_incurred = as.numeric(total_incurred),
    liability_incurred = as.numeric(liability_incurred),
    liability_property_incurred = as.numeric(liability_property_incurred),
    liability_injury_incurred = as.numeric(liability_injury_incurred),
    property_incurred = as.numeric(property_incurred),
    theft_incurred = as.numeric(theft_incurred),
    fire_incurred = as.numeric(fire_incurred),
    glass_incurred = as.numeric(glass_incurred),
    legal_protection_incurred = as.numeric(legal_protection_incurred),
    occupants_incurred = as.numeric(occupants_incurred),
    total_exposure = as.numeric(total_exposure),
    liability_exposure = as.numeric(liability_exposure),
    underwriting_profit = total_premium - total_incurred
    
)

head(df)
```

#identify if the missingness has a pattern.
We observe that the missing values for vehicle values and fuel_type are likely independent.
Whereas vehicle and age are missing together but the value is 2.
If we are using vehicle values and fuel_type it is likely ok to impute for these values.

Missing at Random seems to be in play here.
```{r}
gg_miss_upset(df, nsets = 4)

```



```{r}
vis_dat(df)
vis_miss(df)
```


```{# 1. Pre-filter the data to only include character columns}
df_chr <- df %>% 
  select(where(is.character))

# 2. Create the summary table from the filtered data
unique_summary <- tibble(
  column_name  = names(df_chr),
  unique_count = map_int(df_chr, n_distinct),
  unique_set   = map(df_chr, unique)
)

# 3. View the result
print(unique_summary,n= 40)
```

```{r}
df %>%
  filter(is.na(as.numeric(liability_premium))) %>%
  select(liability_premium) %>%
  distinct()
```


