---
title: "EDA"
format: html
---

```{r}
pacman::p_load(plotly, ggstatsplot, 
               ggdist, ggridges, colorspace,
               gganimate, tidyverse,skimr,naniar,scales)

```


## Read in Data
```{r}
df <-read.csv2("C:/BubblyCoder/ISSS608-VAA/Take_Home_Ex/Dataset of motor insurance portfolio.csv")
head(df)
```


## See overall and missing values.
```{r}

#notice that fuel type has 1287 missing values
# vehicle type has 513 value
# vehicle_age has 2 missing value
# age_driving_license has 2 missing values
skim(df)
```
## see Data types'
```{r}
str(df)
```

#total Exposure max value
motor insurance happen to have only max 1 year contract.
```{r}
max(df$total_exposure)
```


#identify if the missingness has a pattern.
We observe that the missing values for vehicle values and fuel_type are likely independent.
Whereas vehicle and age are missing together but the value is 2.
If we are using vehicle values and fuel_type it is likely ok to impute for these values.

Missing at Random seems to be in play here.
```{r}
gg_miss_upset(df, nsets = 4)

```
## data type conversion
We also created an indicator called underwriting profit to identify which portfolios are losses.
```{r}
df <- df %>% mutate(
    # all the factor like variables
    policy_type = case_when(
      policy_type == "TP"     ~ "Basic Third-Party",
      policy_type == "TPG"    ~ "Third-Party with Glass Cov",
      policy_type == "CC"     ~ "Third-Party with two or more Cov",
      policy_type == "COMP_E" ~ "Comprehensive with Excess",
      policy_type == "COMP_N" ~ "Comprehensive without Excess",
      TRUE                    ~ "Other"
    ),
    policy_status = as.factor(policy_status),
    business_type = as.factor(business_type),
    payment_frequency = as.factor(bonus_score),
    fuel_type = as.factor(fuel_type),
    municipality_type = as.factor(municipality_type),
    circulation_area = as.factor(circulation_area),
    
    # these are all the integer like variables
    vehicle_value = as.numeric(vehicle_value),
    power_to_weight_ratio = as.numeric(power_to_weight_ratio),
    total_premium = as.numeric(total_premium),
    liability_premium = as.numeric(liability_premium),
    property_damage_premium = as.numeric(property_damage_premium),
    theft_premium = as.numeric(theft_premium),
    fire_premium = as.numeric(fire_premium),
    glass_premium = as.numeric(glass_premium),
    legal_protection_premium = as.numeric(legal_protection_premium),
    occupants_premium = as.numeric(occupants_premium),
    total_incurred = as.numeric(total_incurred),
    liability_incurred = as.numeric(liability_incurred),
    liability_property_incurred = as.numeric(liability_property_incurred),
    liability_injury_incurred = as.numeric(liability_injury_incurred),
    property_incurred = as.numeric(property_incurred),
    theft_incurred = as.numeric(theft_incurred),
    fire_incurred = as.numeric(fire_incurred),
    glass_incurred = as.numeric(glass_incurred),
    legal_protection_incurred = as.numeric(legal_protection_incurred),
    occupants_incurred = as.numeric(occupants_incurred),
    total_exposure = as.numeric(total_exposure),
    liability_exposure = as.numeric(liability_exposure),
    underwriting_profit = total_premium - total_incurred
    
) %>% mutate(
  
  policy_type = factor(policy_type, levels = c(
      "Basic Third-Party", 
      "Third-Party with Glass Cov", 
      "Third-Party with two or more Cov", 
      "Comprehensive with Excess", 
      "Comprehensive without Excess",
      "Other"
    ))
  )


head(df)
```

## Low Profitability
A simple way to define this is whatever premium is paid - claims paid out where there are more losses than gains.

Now that we have created an additional column called underwriting profits.
Earlier we also saw that the maximum exposure is 1 year or less in this dataset
```{r}
ggplot(data=df, 
       aes(y = underwriting_profit, x = policy_type)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",        
               size=4) +
  theme(
    # Rotate the x-axis text
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )

```


```{r}
df %>%
  mutate(age_bin = cut(driver_age, breaks = seq(18, 90, by = 10))) %>%
  group_by(policy_description, age_bin) %>%
  summarise(mean_profit = mean(underwriting_profit)) %>%
  ggplot(aes(x = age_bin, y = policy_description, fill = mean_profit)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  labs(title = "Profitability Heatmap",
       fill = "Avg Profit") +
  theme_minimal()
```


