---
title: "Take Home Exercise 2 - Technical Report"
format:
  html:
    code-fold: true
author: "Alvin Ong Wei Ming"
date: "May 15, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: True #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

## Set Up

```{r}
pacman::p_load(jsonlite,dplyr,tidyr,skimr, knitr,kableExtra,ggplot2,stringr,tidyverse,tidygraph,ggraph,patchwork,scales,tidytext,ggalluvial,gt,visNetwork,lubridate)
```

### Read Data

```{r}
data <- fromJSON("data/MC1_final_00.json", simplifyVector = TRUE)
df_rounds_raw <- data$rounds
```

### Unnest Data

```{r}
#| eval: false
# ==============================================================================
# 2. CREATE SEPARATE DATAFRAMES FOR EVERY NESTED STRUCTURE
# ==============================================================================

# --- A. Participants Table ---
# Unnests the participants list and flattens their metadata
df_participants <- df_rounds_raw %>%
  select(hour, participants) %>%
  unnest(participants) %>%
  # unpack is wrapped in a check in case agent_round_metadata is missing in some files
  unpack(any_of("agent_round_metadata"), names_sep = "_")

# --- B. Communications Table ---
# Unnests the chat logs and flattens the agents' internal states
df_communications <- df_rounds_raw %>%
  select(hour, communications) %>%
  unnest(communications) %>%
  unpack(any_of(c("internal_state", "agent_round_metadata")), names_sep = "_")

# --- C. Main Environment & Market Snapshot Table (Scalars only) ---
# Unpacks the environment context, flattens the market snapshot, and removes nested arrays
df_environment <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  unpack(any_of("environment_context_market_snapshot"), names_sep = "_") %>%
  # Select only columns that are NOT lists (this leaves behind the nested arrays for the next step)
  select(where(~ !is.list(.)))

# --- D. Environment Context Nested Arrays ---
# The environment_context contains several arrays (news, media_events, etc.). 
# We unnest them one by one into distinct relational tables linked by 'hour'.

# D1. Media Events
df_media_events <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_media_events) %>%
  unnest(environment_context_media_events, keep_empty = FALSE)

# D2. External Actor Actions
df_external_actions <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_external_actor_actions) %>%
  unnest(environment_context_external_actor_actions, keep_empty = FALSE)

# D3. Social Manager Alerts
df_social_alerts <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_social_manager_alerts) %>%
  unnest(environment_context_social_manager_alerts, keep_empty = FALSE)

# D4. Agents Unavailable
df_agents_unavailable <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_agents_unavailable) %>%
  unnest(environment_context_agents_unavailable, keep_empty = FALSE)

# D5. Critical Deadlines
df_critical_deadlines <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_critical_deadlines) %>%
  unnest(environment_context_critical_deadlines, keep_empty = FALSE)

# D6. News
df_news <- df_rounds_raw %>%
  select(hour, environment_context) %>%
  unpack(environment_context, names_sep = "_") %>%
  select(hour, environment_context_news) %>%
  unnest(environment_context_news, keep_empty = FALSE)
```

### Extract as RData

```{r}
#| eval: false
save(
  df_environment, df_participants, df_communications, 
  df_media_events, df_external_actions, df_social_alerts, 
  df_agents_unavailable, df_critical_deadlines, df_news,
  file = "mc1_unpacked_data.RData"
)
```

## Load data
```{r}
load("mc1_unpacked_data.RData")
```

## Q1 
What were the key events and relationships that led up to the inappropriate release? Visualize the sequence of events. Include elements such as:
	○ Key actions
	○ Causal relationships
	○ Decision points and actors involved in those decisions
Highlight decisions and system elements that were important to the post making it past the embargo enforcement

### Finding the critical moments in Environment Context {#finding-critical-moments}
We start off with a glimpse of the dataset and notice that stock price column has a $ sign, percent change has a % and this prevents it from reading it as a numerical value.
The time stamp is also in chr format and needs to be changed.

```{r}
glimpse(df_environment)
```

```{r}
df_environment <- df_environment %>%
  mutate(
    stock_price_numeric = as.numeric(str_remove(environment_context_market_snapshot_stock_price, "\\$")),
    stock_percent_numeric = as.numeric(str_remove(environment_context_market_snapshot_percent_change, "\\%")),

    hour_datetime = as.POSIXct(hour, format = "%Y-%m-%dT%H:%M:%S")
  ) %>%
  # CRITICAL STEP: Arrange rows in perfect chronological order
  arrange(hour_datetime)
glimpse(df_environment)
```


```{r}

# 1. Prepare and cleanly sort the data
df_plot_clean <- df_environment %>%
  mutate(
    # Strip the '$' and convert to a clean decimal number
    stock_price_numeric = as.numeric(str_remove(environment_context_market_snapshot_stock_price, "\\$")),
    
    # Force convert 'hour' string into a real, raw POSIXct datetime object
    hour_datetime = as.POSIXct(hour, format = "%Y-%m-%dT%H:%M:%S")
  ) %>%
  # Drop the hours where the stock price is missing (the 5 Nulls)
  filter(!is.na(stock_price_numeric)) %>%
  # CRITICAL STEP: Arrange rows in perfect chronological order
  arrange(hour_datetime)

# 2. Plot using the datetime column (We drop 'group = 1' because it's continuous now!)
ggplot(df_plot_clean, aes(x = hour_datetime, y = stock_price_numeric)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "darkblue", size = 1.5) + 
  # This dynamically formats the timeline ticks on the X-axis nicely
  scale_x_datetime(date_labels = "%b %d\n%H:%M", date_breaks = "2 days") +
  theme_minimal() +
  labs(
    title = "TenantThread Stock Price Leading Up to Crisis",
    subtitle = "Cleaned chronological timeline view",
    x = "Date & Time",
    y = "Stock Price ($)"
  )
```
We can quickly observed from the above graph that the stock price started to spike up on 4th June and this is likely where our starting point will be
to understanding what happen prior to the stock price going up.

```{r}
june_4th_clues <- df_environment %>%
  filter(str_detect(hour, "2046-06-04")) %>%
  select(hour, environment_context_event_narrative,environment_context_event_headline, environment_context_social_state)

knitr::kable(june_4th_clues, format = "markdown")
```

We also observed the hashtag  #AlgorithmicEviction here that is align with the background so far.
The sentiment deteoriating rapidly also suggest we should explore the other available data on what news triggered this drop

## Investigate News data

```{r}
knitr::kable(df_news %>%
  filter(str_detect(hour, "2046-06-04")))
```

There is already mention of "Crsis" in 4th June. lets investigate further onto the other available data.
We now look into communications among the agents to find out what can we observed that led to this.

We apply TD-IDF below to get a sensing of what each day happen without going through each text

Setting colors for the various agent for easy identification

### Color palette
```{r}

#| code-fold: False
agent_palette <- c(
  "judge_agent"        = "#b38600",  # Dark Gold / Yellow
  "legal_agent"        = "#4daf4a",  # Corporate Green
  "pr_agent"           = "#00a699",  # Dark Teal
  "social_media_agent" = "#e377c2",  # Crisis Pink
  "intern_agent"       = "#f07167",  # Coral / Light Red
  "pr_intern_agent"    = "#1f77b4",  # Cool Blue
  "quality_agent"      = "#9467bd"   # Infrastructure Purple
)
```


::: panel-tabset
### Running on 4th and 5th June
```{r}
#| fig-width: 16
#| fig-height: 8
# 1. Tokenize and clean the internal deliberation text
df_words <- df_communications %>%
  filter(str_detect(hour, "2046-06-04|2046-06-05")) %>%
  mutate(
    # Combine internal states and message contents
    full_text = paste(replace_na(internal_state_deliberating, ""), 
                      replace_na(content, ""))
  ) %>%
  select(hour, agent_id, full_text) %>%
  # Break text down into individual words
  unnest_tokens(word, full_text) %>%
  # Remove standard stop words (the, and, of, etc.)
  anti_join(get_stopwords(), by = "word") %>%
  # Filter out numbers and punctuation artifacts
  filter(str_detect(word, "^[a-zA-Z]+$"))

# 2. Compute TF-IDF to find unique themes per Agent during the crisis
df_tf_idf <- df_words %>%
  count(agent_id, word, sort = TRUE) %>%
  bind_tf_idf(word, agent_id, n)

# 3. Plot the top distinct topic terms for each agent
df_tf_idf %>%
  group_by(agent_id) %>%
  slice_max(tf_idf, n = 7, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, tf_idf, agent_id)) %>%
  ggplot(aes(x = tf_idf, y = word, fill = agent_id)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  scale_fill_manual(values = agent_palette) +
  facet_wrap(~ agent_id, scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(axis.text.y = element_text(face = "bold")) +
  labs(
    title = "Structural Text Analysis: Core Focus Areas per Agent (4th June & 5th June)",
    subtitle = "Top unique terms extracted via TF-IDF reveal distinct operational priorities during the crisis.",
    x = "Theme Importance (TF-IDF Score)",
    y = NULL
  )

```

### Running on 4th

```{r}
#| fig-width: 16
#| fig-height: 8
# 1. Tokenize and clean the internal deliberation text
df_words <- df_communications %>%
  filter(str_detect(hour, "2046-06-04")) %>%
  mutate(
    # Combine internal states and message contents
    full_text = paste(replace_na(internal_state_deliberating, ""), 
                      replace_na(content, ""))
  ) %>%
  select(hour, agent_id, full_text) %>%
  # Break text down into individual words
  unnest_tokens(word, full_text) %>%
  # Remove standard stop words (the, and, of, etc.)
  anti_join(get_stopwords(), by = "word") %>%
  # Filter out numbers and punctuation artifacts
  filter(str_detect(word, "^[a-zA-Z]+$"))

# 2. Compute TF-IDF to find unique themes per Agent during the crisis
df_tf_idf <- df_words %>%
  count(agent_id, word, sort = TRUE) %>%
  bind_tf_idf(word, agent_id, n)

# 3. Plot the top distinct topic terms for each agent
df_tf_idf %>%
  group_by(agent_id) %>%
  slice_max(tf_idf, n = 7, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, tf_idf, agent_id)) %>%
  ggplot(aes(x = tf_idf, y = word, fill = agent_id)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  scale_fill_manual(values = agent_palette) +
  facet_wrap(~ agent_id, scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(axis.text.y = element_text(face = "bold")) +
  labs(
    title = "Structural Text Analysis: Core Focus Areas per Agent (4th June)",
    subtitle = "Top unique terms extracted via TF-IDF reveal distinct operational priorities during the crisis.",
    x = "Theme Importance (TF-IDF Score)",
    y = NULL
  )
```

### Running on 5th only

```{r}
#| fig-width: 16
#| fig-height: 8
# 1. Tokenize and clean the internal deliberation text
df_words <- df_communications %>%
  filter(str_detect(hour, "2046-06-05")) %>%
  mutate(
    # Combine internal states and message contents
    full_text = paste(replace_na(internal_state_deliberating, ""), 
                      replace_na(content, ""))
  ) %>%
  select(hour, agent_id, full_text) %>%
  # Break text down into individual words
  unnest_tokens(word, full_text) %>%
  # Remove standard stop words (the, and, of, etc.)
  anti_join(get_stopwords(), by = "word") %>%
  # Filter out numbers and punctuation artifacts
  filter(str_detect(word, "^[a-zA-Z]+$"))

# 2. Compute TF-IDF to find unique themes per Agent during the crisis
df_tf_idf <- df_words %>%
  count(agent_id, word, sort = TRUE) %>%
  bind_tf_idf(word, agent_id, n)

# 3. Plot the top distinct topic terms for each agent
df_tf_idf %>%
  group_by(agent_id) %>%
  slice_max(tf_idf, n = 7, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, tf_idf, agent_id)) %>%
  ggplot(aes(x = tf_idf, y = word, fill = agent_id)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  scale_fill_manual(values = agent_palette) +
  facet_wrap(~ agent_id, scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(axis.text.y = element_text(face = "bold")) +
  labs(
    title = "Structural Text Analysis: Core Focus Areas per Agent (5tH June)",
    subtitle = "Top unique terms extracted via TF-IDF reveal distinct operational priorities during the crisis.",
    x = "Theme Importance (TF-IDF Score)",
    y = NULL
  )
```
:::
When we look through each graph, we start to notice that:

- Judge agent has a that theme appears to be restraining itself from words such as compliance, hold, truth, flag

- PR agent appears to be writing documents with negativity and carefulness with words  such as objection, exclusive, selective, ethical

- Quality agents appears to be going through many briefings with others with words such as walkthrough, memo, briefings, board and Ajay the ceo of the company

- Social Media agent appears to be doing simulation and words with negativity appears such as denial, critical, projected. There is also mention of the share and appears to be simulation of share price.

- PR intern appears to be doing analysis on data on rent.

- Intern agent appears to be unware of what is going on and has similar theme throughout the 3 days.






### Alluvial Chart
Although we have a clue from the TF-IDF Study on the theme from 4th June to 5th June,we would like to dive deeper to understand
which agent had the most involvement in the change in stock price or perception of company after the news after the merger leaked out.

```{r}
# 1. Categorize timeline and establish forensic priority weights
df_alluvial_prep <- df_communications %>%
  mutate(
    phase = case_when(
      str_detect(hour, "2046-06-04") ~ "1. June 4th",
      str_detect(hour, "2046-06-05T09|2046-06-05T10|2046-06-05T11|2046-06-05T12|2046-06-05T13|2046-06-05T14|2046-06-05T15") ~ "2. June 5th\n(9am - 3pm)",
      str_detect(hour, "2046-06-05T16") ~ "3. June 5th \n(4pm - 5pm)",
      str_detect(hour, "2046-06-05T17") ~ "4. June 5th 5:12 PM\n(Breach)",
      TRUE ~ NA_character_
    ),
    channel_priority = case_when(
      channel %in% c("personal_post", "anonymous_post") ~ 4,  
      channel %in% c("side_huddle", "one_on_one_chat")  ~ 3,  
      channel == "comms_huddle"                         ~ 1,  
      TRUE                                              ~ 2   
    )
  ) %>%
  filter(!is.na(phase)) %>%
  group_by(phase, channel, agent_id, channel_priority) %>%
  summarise(msg_volume = n(), .groups = "drop") %>%
  
  group_by(phase, agent_id) %>%
  slice_max(channel_priority, n = 1, with_ties = FALSE) %>%
  ungroup()

df_alluvial_prep$phase <- factor(df_alluvial_prep$phase, 
                                 levels = c("1. June 4th", 
                                            "2. June 5th\n(9am - 3pm)", 
                                            "3. June 5th \n(4pm - 5pm)", 
                                            "4. June 5th 5:12 PM\n(Breach)"))


ggplot(df_alluvial_prep, aes(x = phase, y = msg_volume, stratum = channel, alluvium = agent_id, fill = agent_id)) +
  geom_alluvium(alpha = 0.65, width = 1/4, color = "white", linewidth = 0.1) +
  geom_stratum(alpha = 0.15, width = 1/4, fill = "black", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), fontface = "bold", size = 3) +
  
  scale_fill_manual(values = agent_palette) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold", size = 11, color = "black"), # Fixed linewidth to size
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  labs(
    title = "System Interaction Alluvial Story Tracker: The Path to the Breach",
    subtitle = "Tracking how individual agent communications migrated across operational environments as the crisis escalated.",
    x = "Investigation Timeline Milestones",
    y = "Operational Traffic Mass",
    fill = "Executing AI Actor"
  )
```

This chart uncovered that the judge agent had no communications after 3pm although its the compliance enforcer.
After 3pm, the legal agent and pr intern agent continued communications with one on one chat leading to an anonymous post.
Whereas the social medial agent also continued with a personal post.
Hence we now narrow down the 3 agents that we should focus on.


```{r}
#| column: page
#| warning: false
knitr::kable(df_communications %>% filter(str_detect(hour, "2046-06-05") & agent_id %in% c('legal_agent','pr_intern_agent','social_media_agent')) %>% select(agent_id, timestamp,internal_state_deliberating,content) )
```


### Word Occurrence Network Graph  
Next we try to use a Bi-Gram approach to create a word occurence graph to try to narrow down 
what the 3 agents are saying prior to the release of the embargo to get a sensing what happened.

```{r}
#| fig-width: 16
#| fig-height: 8

df_bigrams <- df_communications %>%
  filter(str_detect(hour, "2046-06-05T16|2046-06-05T17"), agent_id %in% c('legal_agent','pr_intern_agent','social_media_agent')) %>% # Focus on the crisis day
  mutate(
    # Clean and concatenate both target columns
    full_text = paste(replace_na(internal_state_deliberating, ""), 
                      replace_na(content, ""))
  ) %>%
  select(full_text) %>%
  unnest_tokens(bigram, full_text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  # Remove standard stop words from both positions
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         str_detect(word1, "^[a-zA-Z]+$"),
         str_detect(word2, "^[a-zA-Z]+$")) %>%
  # Count occurrences to find the heavy-hitting phrases
  count(word1, word2, sort = TRUE) %>%
  # Filter out rare pairings to keep the network graph clean
  slice_max(n, n = 70, with_ties = FALSE) 

# 2. Convert to a directional graph object
graph_bigrams <- as_tbl_graph(df_bigrams, directed = TRUE)

# 3. Plot the Phrase Network
ggraph(graph_bigrams, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), 
                 color = "steelblue",
                 arrow = arrow(length = unit(2.5, 'mm')), 
                 end_cap = circle(2.5, 'mm')) +
  geom_node_point(color = "grey20", size = 1.5) +
  geom_node_text(aes(label = name), repel = TRUE, size = 4, fontface = "bold") +
  scale_edge_width(range = c(0.4, 2)) +
  theme_graph() +
  labs(
    title = "Phrase Architecture: Word Co-occurrence Network (June 5th)",
    subtitle = "Links uncover actionable word couplings inside internal thoughts and outgoing messages.",
    edge_width = "Phrase Frequency",
    edge_alpha = "Phrase Frequency"
  )
```

It doesnt feel too intuitive from the above.

```{r}


df_agent_phrases <- df_communications %>%
  filter(
    str_detect(hour, "2046-06-05T16|2046-06-05T17"), 
    agent_id %in% c('legal_agent', 'pr_intern_agent', 'social_media_agent')
  ) %>%
  mutate(
    full_text = paste(replace_na(internal_state_deliberating, ""), 
                      replace_na(content, ""))
  ) %>%
  select(agent_id, full_text) %>%
  unnest_tokens(bigram, full_text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         str_detect(word1, "^[a-zA-Z]+$"),
         str_detect(word2, "^[a-zA-Z]+$")) %>%
  mutate(clean_phrase = paste(word1, word2, sep = " ")) %>%
  

  count(agent_id, clean_phrase, sort = TRUE) %>%
  group_by(agent_id) %>%
  slice_max(n, n = 6, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(clean_phrase = reorder_within(clean_phrase, n, agent_id))

ggplot(df_agent_phrases, aes(x = n, y = clean_phrase, fill = agent_id)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  scale_fill_manual(values = agent_palette) +
  facet_wrap(~ agent_id, scales = "free_y", ncol = 1) + # Stacked panels for direct comparison
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold", size = 12),
    axis.text.y = element_text(face = "bold", size = 10)
  ) +
  labs(
    title = "Dialogue Focus by Agent",
    subtitle = "Isolating the top distinct phrases used by each actor during the breakout hour.",
    x = "Phrase Frequency Count",
    y = NULL
  )
```
We found that all 3 agents top phrase had "press release" and "bilateral consent".
The official flex post is also present in the legal agent and pr intern agent. 
These keyswords appears to guide us to the truth.
We shall now attempt to filter for these keywords 
Note: 36 rows found

### Full Relevant Content
```{r}


df_communications %>% 
  filter(
    str_detect(hour, "2046-06-05T17"),
    str_detect(content, "flex|consent|release|critical") | 
    str_detect(internal_state_deliberating, "flex|consent|release|critical")
  ) %>% 
  select(agent_id, channel, content, internal_state_deliberating, hour) %>% 
  
  # ==============================================================================
  # THE FORENSIC SANITIZER: Clean text columns to prevent HTML tag rupture
  # ==============================================================================
  mutate(
    # 1. Strip out raw carriage returns and newlines
    content = str_replace_all(content, "[\r\n]+", " "),
    internal_state_deliberating = str_replace_all(internal_state_deliberating, "[\r\n]+", " "),
    
    # 2. Optional: Cap length at 400 characters so columns don't stretch forever
    content = str_trunc(content, width = 400, side = "right"),
    internal_state_deliberating = str_trunc(internal_state_deliberating, width = 400, side = "right")
  ) %>%
  
  # ==============================================================================
  # RENDER CLEAN CONTAINER TABLE
  # ==============================================================================
  knitr::kable(format = "html") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  kableExtra::scroll_box(
    height = "400px",      
    width = "100%",        
    fixed_thead = TRUE     
  )
```

From the above filtered conversation, we can observe that the post is intentionally release by the agents.
Consent was received from CivicLoom's legal team hence "bilateral consent" kept appearing in the previous. The release was performed by pr intern and the judge also made sure compliance was enforced with "Judge — CivicLoom verbal consent confirmed. Section 4.3(c) bilateral release effective immediately. Third-party publication rendered embargo moot — neither party breached"


## Q2 
The evasion of the embargo was a new behavior. Provide visual analytics that discover and illustrate the typical behavior of systems/agents/people involved. How does the behavior that led to the release compare to the prior behaviors?

We learnt earlier from [Finding the critical moments](#finding-critical-moments) that the critical point of the change was on '2046-06-04', hence we will explore the normal behaviors of the agents before the critical date.


### Typical Behavior

```{r}
#| warning: false
#| message: false
#| layout-ncol: 2
#| column: page


custom_stop_words <- tibble(word = c("legal", "agent", "pr", "intern", "social", "media", "judge", "quality"))

# 2. Extract baseline text data (Up to June 3rd)
baseline_words <- df_communications %>%
  filter(hour < "2046-06-04") %>%
  mutate(full_text = paste(replace_na(internal_state_deliberating, ""), replace_na(content, ""))) %>%
  select(agent_id, full_text) %>%
  unnest_tokens(word, full_text) %>%
  anti_join(get_stopwords(), by = "word") %>%
  anti_join(custom_stop_words, by = "word") %>%
  filter(str_detect(word, "^[a-zA-Z]+$")) %>%
  count(agent_id, word) %>%
  bind_tf_idf(word, agent_id, n) %>%
  group_by(agent_id) %>%
  slice_max(tf_idf, n = 10, with_ties = FALSE) %>%
  ungroup()

# 3. Create Edges Data Frame for visNetwork
edges_baseline <- baseline_words %>%
  rename(from = agent_id, to = word, value = tf_idf)

# 4. Create Nodes Data Frame with explicit mapping attributes
nodes_baseline <- tibble(name = unique(c(edges_baseline$from, edges_baseline$to))) %>%
  mutate(
    id = name,
    label = name,
    group = if_else(name %in% edges_baseline$from, "Agent", "Keyword"),
    # Assign node sizes dynamically
    size = if_else(group == "Agent", 35, 15),
    # Inject specific palette mapping values
    color = if_else(group == "Agent", agent_palette[name], "#9ecae1")
    )


# 5. Render Interactive Baseline Network
p1 <- visNetwork(nodes_baseline, edges_baseline, 
                 main = "Baseline Network Structure",
                 submain = "18th May  - 3rd June  Typical Behavior",
                 width = "100%", height = "650px") %>%
  visGroups(groupname = "Agent", font = list(size = 16, face = "bold")) %>%
  visGroups(groupname = "Keyword", font = list(size = 12)) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
             nodesIdSelection = TRUE) %>%
  visPhysics(solver = "forceAtlas2Based", stabilization = TRUE) %>%
  visLayout(randomSeed = 123)




# 1. Extract crisis text data (June 4th & 5th)
crisis_words <- df_communications %>%
  filter(hour >= "2046-06-04") %>%
  mutate(full_text = paste(replace_na(internal_state_deliberating, ""), replace_na(content, ""))) %>%
  select(agent_id, full_text) %>%
  unnest_tokens(word, full_text) %>%
  anti_join(get_stopwords(), by = "word") %>%
  anti_join(custom_stop_words, by = "word") %>%
  filter(str_detect(word, "^[a-zA-Z]+$")) %>%
  count(agent_id, word) %>%
  bind_tf_idf(word, agent_id, n) %>%
  group_by(agent_id) %>%
  slice_max(tf_idf, n = 10, with_ties = FALSE) %>%
  ungroup()

# 2. Create Edges Data Frame
edges_crisis <- crisis_words %>%
  rename(from = agent_id, to = word, value = tf_idf)

# 3. Create Nodes Data Frame
nodes_crisis <- tibble(name = unique(c(edges_crisis$from, edges_crisis$to))) %>%
  mutate(
    id = name,
    label = name,
    group = if_else(name %in% edges_crisis$from, "Agent", "Keyword"),
    size = if_else(group == "Agent", 35, 15),
    color = if_else(group == "Agent", agent_palette[name], "#9ecae1")
  )

# 4. Render Interactive Crisis Network
p2 <- visNetwork(nodes_crisis, edges_crisis,
                 main = "Crisis Network Structure",
                 submain = "4th June onwards Typical Behavior",width = "100%", height = "650px") %>%
  visGroups(groupname = "Agent", font = list(size = 16, face = "bold")) %>%
  visGroups(groupname = "Keyword", font = list(size = 12)) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
             nodesIdSelection = TRUE) %>%
  visPhysics(solver = "forceAtlas2Based", stabilization = TRUE) %>%
  visLayout(randomSeed = 123)

p1
p2
```

To interpret the above chart, the thicker the line the more unique the word is to the agent through the TF IDF transformation.

On the left (Typical behavior), most agents operate within isolated star-burst formations, communicating via distinct, non-overlapping vocabularies tailored to their explicit job roles. Distinctively, the quality_agent and legal_agent form a heavily weighted structural bridge through the keyword risk, signifying routine corporate governance, auditing, and compliance monitoring. Aside from a minor conversational convergence around the filler word us among the legal and public relations agents, key executioners like the pr_intern_agent and intern_agent remain completely disconnected from the rest of the corporate infrastructure, working in total isolation.

On the right (criss behavior), we can see structural shifts in the network.
A new, targeted coordination bridge emerges directly between the legal_agent and social_media_agent through the shared keyword pm, this likely captures capturing a synchronized effort to manage the precise timing of the public release.
We also notice that intern_agent and pr_intern_agent suddenly link through a heavily weighted focus on morale.

## Q3
3. Were there leading indicators that such a release was possible?
- Are there prior occasions where agent’s actual behaviors differ from their expected behavior? When? Which behaviors were observed?
- Were there other occasions where the agent system exhibited behaviors like those that led to the release? When? Which behaviors were observed?
Why didn’t the prior occasions result in noticeable action?

```{r}
#| fig-width: 14
#| fig-height: 9
#| column: page
#| out-width: "100%"

df_heatmap_prep <- df_communications %>%
  mutate(
    # Parse ISO8601 strings (e.g., "2046-06-05T17") safely
    parsed_time = as_datetime(hour),
    date_label  = as.Date(parsed_time),
    hour_of_day = hour(parsed_time)
  ) %>%
  # Count total message volume per agent, per day, per hour
  count(agent_id, date_label, hour_of_day, name = "message_count")

# 2. Build the high-contrast grid matrix
ggplot(df_heatmap_prep, aes(x = hour_of_day, y = factor(date_label), fill = message_count)) +
  geom_tile(color = "white", size = 0.1) +
  # Magma palette makes structural anomalies glow brightly against dark baselines
  scale_fill_viridis_c(option = "magma", name = "Log Volume") +
  # Separate grids for each agent with independent y-axes to accommodate active ranges
  facet_wrap(~ agent_id, scales = "free_y", ncol = 2) +
  theme_minimal(base_family = "Arial") +
  theme(
    strip.text = element_text(face = "bold", size  = 11),
    axis.text = element_text(size = 9),
    panel.spacing = unit(1, "lines"),
    legend.position = "bottom"
  ) +
  labs(
    title = "System Interaction Rhythms: Operational Heatmap Matrix",
    subtitle = "Identifying out-of-bounds behavioral anomalies and off-hour communication spikes by actor.",
    x = "Hour of Day (24h format)",
    y = "Operational Date"
  )
```

