---
title: "Hands-on_Ex05A"
author: "Alvin Ong Wei Ming"
date: "May 10, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: true #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

# Creating Ternary Plot with R

What is a ternary plot?

Ternary plots are a way of displaying the distribution and variability of three-part compositional data. (For example, the proportion of aged, economy active and young population or sand, silt, and clay in soil.) It’s display is a triangle with sides scaled from 0 to 1. Each side represents one of the three components. A point is plotted so that a line drawn perpendicular from the point to each leg of the triangle intersect at the component values of the point.

For this exercise we will use the below 2 packages

-   [ggtern](https://www.rdocumentation.org/packages/ggtern/versions/4.0.0)a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.

-   [Plotly R](https://plotly.com/r/) an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library, plotly.js . The plotly R libary contains the ggplotly function, which will convert ggplot2 figures into a Plotly object.

```{r}
pacman::p_load(plotly, ggtern, tidyverse)
```

We will use the Singapore Residents by Planning AreaSubzone, Age Group, Sex and Type of Dwelling, June 2000-2018 data

```{r}
#Reading the data into R environment
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv")
head(pop_data)
```

## Preparing the data

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)

head(agpop_mutated)
```

::: callout-note
mutate: To create a new column spread: It takes the age groups (stored in AG) and turns each unique age group into its own column, filling the cells with the corresponding Population value
:::

## Plotting Ternary Diagram with R

### Plotting a static ternary diagram

We use ggtern to create a simple ternary plot.

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()

```

Upgrading the plot with colours.

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

We can also add kernel density estimate across the three variables and draws "contour lines" where the data points are most concentrated.

```{r}
ggtern(data = agpop_mutated, aes(x = YOUNG, y = ACTIVE, z = OLD)) +
  geom_point(alpha = 0.5) + # Keep points slightly transparent
  geom_density_tern(color = "darkred", 
                    linewidth = 0.5, 
                    linetype = "dashed") + 
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

## Plotting an interative ternary diagram

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )

```
