---
title: "Hands-on_Ex05E"
format: html
---


## Data Wrangling

```{r}
realis2018 <- read_csv("data/realis2018.csv")
head(realis2018)
```

The data.frame realis2018 is in trasaction record form, which is highly disaggregated and not appropriate to be used to plot a treemap. In this section, we will perform the following steps to manipulate and prepare a data.frtame that is appropriate for treemap visualisation:

group transaction records by Project Name, Planning Region, Planning Area, Property Type and Type of Sale, and
compute Total Unit Sold, Total Area, Median Unit Price and Median Transacted Price by applying appropriate summary statistics on No. of Units, Area (sqm), Unit Price ($ psm) and Transacted Price ($) respectively.
Two key verbs of dplyr package, namely: group_by() and summarize() will be used to perform these steps.

group_by() breaks down a data.frame into specified groups of rows. When you then apply the verbs above on the resulting object they’ll be automatically applied “by group”.

Grouping affects the verbs as follows:

grouped select() is the same as ungrouped select(), except that grouping variables are always retained.
grouped arrange() is the same as ungrouped; unless you set .by_group = TRUE, in which case it orders first by the grouping variables.
mutate() and filter() are most useful in conjunction with window functions (like rank(), or min(x) == x). They are described in detail in vignette(“window-functions”).
sample_n() and sample_frac() sample the specified number/fraction of rows in each group.
summarise() computes the summary for each group.
In our case, group_by() will used together with summarise() to derive the summarised data.frame.

## Grouped summaries without the Pipe


```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))

```

::: {.callout-note}
Aggregation functions such as sum() and meadian() obey the usual rule of missing values: if there’s any missing value in the input, the output will be a missing value. The argument na.rm = TRUE removes the missing values prior to computation.
:::

## Grouped summaries with the pipe
The code chunk below shows a more efficient way to tackle the same processes by using the pipe, %>%:

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```
---
