---
title: "Hands-on_Ex05B"
author: "Alvin Ong Wei Ming"
date: "May 10, 2026"
date-modified: "last-modified"
execute:
  eval: true # As and when you render the code it will run through and display the output
  echo: true #when it render the document it will show the code chunk
  warning: false # this will prevent warning from appearing.
  freeze: false #everytime you render, it will only render my latest hands on exericse and not the rest of it.
---

# Visual Correlation Analysis

Correlation coefficient is a popular statistic that use to measure the type and strength of the relationship between two variables. The values of a correlation coefficient ranges between -1.0 and 1.0. A correlation coefficient of 1 shows a perfect linear relationship between the two variables, while a -1.0 shows a perfect inverse relationship between the two variables. A correlation coefficient of 0.0 shows no linear relationship between the two variables.

When multivariate data are used, the correlation coefficeints of the pair comparisons are displayed in a table form known as correlation matrix or scatterplot matrix.

There are three broad reasons for computing a correlation matrix.

-   To reveal the relationship between high-dimensional variables pair-wisely.

-   To input into other analyses. For example, people commonly use correlation matrices as inputs for exploratory factor analysis, confirmatory factor analysis, structural equation models, and linear regression when excluding missing values pairwise.

-   As a diagnostic when checking other analyses. For example, with linear regression a high amount of correlations suggests that the linear regression’s estimates will be unreliable.

When the data is large, both in terms of the number of observations and the number of variables, [Corrgram](https://www.datavis.ca/papers/corrgram.pdf) tend to be used to visually explore and analyse the structure and the patterns of relations among variables. It is designed based on two main schemes:

-   Rendering the value of a correlation to depict its sign and magnitude, and

-   Reordering the variables in a correlation matrix so that “similar” variables are positioned adjacently, facilitating perception.

::: callout-tip
The [Corrgram](https://www.datavis.ca/papers/corrgram.pdf) is a good read, learning about partial correlation charts and its importance in scenarios to dispel Simpon's paradox
:::

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)

```

## The Dataset

In this hands-on exercise, the Wine Quality Data Set of UCI Machine Learning Repository will be used. The data set consists of 13 variables and 6497 observations. For the purpose of this exercise, we have combined the red wine and white wine data into one data file. It is called wine_quality and is in csv file format.

```{r}
wine <- read_csv("data/wine_quality.csv")
head(wine)
```

## Building Correlation Matrix: pairs() method

Figure below shows the scatter plot matrix of Wine Quality Data. It is a 11 by 11 matrix.

```{r}
pairs(wine[,1:11])

```

The required input of pairs() can be a matrix or data frame. The code chunk used to create the scatterplot matrix is relatively simple. It uses the default pairs function. Columns 2 to 12 of wine dataframe is used to build the scatterplot matrix. The variables are: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates and alcohol.

```{r}
pairs(wine[2:12])

```

## To show lower panel only

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

## to show upper panel only

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

## Including with correlation coefficients

To show the correlation coefficient of each pair of variables instead of a scatter plot, panel.cor function will be used. This will also show higher correlations in a larger font.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr=usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

## Visualising Correlation Matrix: ggcormat()

One of the major limitation of the correlation matrix is that the scatter plots appear very cluttered when the number of observations is relatively large (i.e. more than 500 observations). To over come this problem, Corrgram data visualisation technique suggested by D. J. Murdoch and E. D. Chow (1996) and Friendly, M (2002) and will be used.

The are at least three R packages provide function to plot corrgram, they are:

-   [corrgram](https://cran.r-project.org/web/packages/corrgram/index.html)
-   [ellipse](https://cran.r-project.org/web/packages/ellipse/index.html)
-   [corrplot](https://cran.r-project.org/web/packages/corrplot/index.html)

On top that, some R package like ggstatsplot package also provides functions for building corrgram.

## The basic plot

```{r}
# Feels similar to seaborn heatmap
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 0.7),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are not significant at p < 0.05"
)

```

Things to learn from the code chunk above:

-   cor.vars argument is used to compute the correlation matrix needed to build the corrgram.

-   ggcorrplot.args argument provide additional (mostly aesthetic) arguments that will be passed to ggcorrplot::ggcorrplot function. The list should avoid any of the following arguments since they are already internally being used: corr, method, p.mat, sig.level, ggtheme, colors, lab, pch, legend.title, digits.

The sample sub-code chunk can be used to control specific component of the plot such as the font size of the x-axis, y-axis, and the statistical report

ggplot.component = list( theme(text=element_text(size=5), axis.text.x = element_text(size = 8), axis.text.y = element_text(size = 8)))

::: callout-note
-   hc.order = True refers to using hierarchical clustering to put similar variables together while tl.cex = 10: Adjusts the text label size for the variable names so they are readable even if the matrix is large.

-   The current default uses p value, a bayesian approach can also be used by changing type = 'bayes'
:::

```{r}

#Changing the color to better show that this is different from the above plots

ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  type = "bayes",
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 0.7),
  title    = "Bayesian Correlogram for Wine Dataset",
  subtitle = "Values in cells represent Log(Bayes Factor)"
) + scale_fill_distiller(palette = "YlOrRd", direction = 1)
```

After switching from frequentist to bayes, none of the plot was crossed out.

A quick search: While the frequentist approach suggests four pairs are non-significant, the Bayesian approach reveals that these relationships are 'anecdotal' ($1 < BF_{10} < 3$) rather than definitively absent, providing a more nuanced view of the chemical interactions in the wine

Some examples of interpretation

$log_e(BF_{10})$: This is the "Log-Bayes Factor." We use logs because Bayes Factors can become massive (e.g., 1,000,000).Positive number (e.g., 5): Strong evidence for a relationship.Negative number (e.g., -2): Evidence against a relationship (supports the null).Near 0: The data is "inconclusive" or "anecdotal."

$BF_{10} = 10$: The data is 10 times more likely under the "Relationship" model than the "No Relationship" model.

## Building multiple plots

Since ggstasplot is an extension of ggplot2, it also supports faceting. However the feature is not available in ggcorrmat() but in the [grouped_ggcorrmat()](https://www.indrapatil.com/ggstatsplot/reference/grouped_ggcorrmat.html) of ggstatsplot.

```{r}

#| fig-width: 32    # 1. Make the canvas very wide
#| fig-height: 16    # 1. Make the canvas taller
#| out-width: "100%"
#| 


grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 0.6,
                         tl.srt= 45
                         ),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

Things to learn from the code chunk above:

-   to build a facet plot, the only argument needed is grouping.var.
-   Behind group_ggcorrmat(), patchwork package is used to create the multiplot. plotgrid.args argument provides a list of additional arguments passed to patchwork::wrap_plots, except for guides argument which is already separately specified earlier.
-   Likewise, annotation.args argument is calling plot annotation arguments of patchwork package.

## Visualising Correlation Matrix using corrplot Package

Here is an introduction to An Introduction to [corrplot Package](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

This package is amazing !

::: callout-tip
Did you know [corrplot](https://pypi.org/project/ggcorrplot/#doc) is also available in python. Kudos to the team for exporting this amazing package to python too But its still lacks the flexibility that is available in R.

Incidentally found the link to [Altair](https://altair-viz.github.io/gallery/isotype_emoji.html). I just need to find the equivalent for R for this !
:::

## Before we can plot a corrgram using corrplot(), we need to compute the correlation matrix of wine data frame.

In the code chunk below, cor() of R Stats is used to compute the correlation matrix of wine data frame.

```{r}
wine.cor <- cor(wine[, 1:11])
corrplot(wine.cor)
```

Notice that the default visual object used to plot the corrgram is circle. The default layout of the corrgram is a symmetric matrix. The default colour scheme is diverging blue-red. Blue colours are used to represent pair variables with positive correlation coefficients and red colours are used to represent pair variables with negative correlation coefficients. The intensity of the colour or also know as saturation is used to represent the strength of the correlation coefficient. Darker colours indicate relatively stronger linear relationship between the paired variables. On the other hand, lighter colours indicates relatively weaker linear relationship.

## Working with visual geometrics

In corrplot package, there are seven visual geometrics (parameter method) can be used to encode the attribute values. They are: circle, square, ellipse, number, shade, color and pie. The default is circle. As shown in the previous section, the default visual geometric of corrplot matrix is circle. However, this default setting can be changed by using the method argument as shown in the code chunk below

::: panel-tabset
## ellipse

```{r}
corrplot(wine.cor, 
         method = "ellipse") 
```

### square

```{r}
corrplot(wine.cor, 
         method = "square") 
```

## hclust

```{r}
## If using 'hclust', corrplot() can draw rectangles around the plot of correlation matrix based on the results of hierarchical clustering.
## FIX: Use corrplot() function, not the variable name
corrplot(wine.cor, order = 'hclust', addrect = 2)

```

## hist mixed

```{r}

# 3. Create the Mixed Correlogram
corrplot.mixed(wine.cor, 
               lower = 'shade', 
               upper = 'pie', 
               order = 'hclust',
               tl.col = "black", 
               tl.cex = 0.7,
               addrect = 3)     # Optional: Draws boxes around hclust groups
```
:::

## Working with layout

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")


```

The default layout of the corrgram can be further customised. For example, arguments diag and tl.col are used to turn off the diagonal cells and to change the axis text label colour to black colour respectively as shown in the code chunk and figure below

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")

```

## Working with mixed layout

With corrplot package, it is possible to design corrgram with mixed visual matrix of one half and numerical matrix on the other half. In order to create a coorgram with mixed layout, the corrplot.mixed(), a wrapped function for mixed visualisation style will be used.

Figure below shows a mixed layout corrgram plotted using wine quality data.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

Notice that argument lower and upper are used to define the visualisation method used. In this case ellipse is used to map the lower half of the corrgram and numerical matrix (i.e. number) is used to map the upper half of the corrgram. The argument tl.pos, on the other, is used to specify the placement of the axis label. Lastly, the diag argument is used to specify the glyph on the principal diagonal of the corrgram.

## Combining corrgram with the significant test

In statistical analysis, we are also interested to know which pair of variables their correlation coefficients are statistically significant.

Figure below shows a corrgram combined with the significant test. The corrgram reveals that not all correlation pairs are statistically significant. For example the correlation between total sulfur dioxide and free surfur dioxide is statistically significant at significant level of 0.1 but not the pair between total sulfur dioxide and citric acid.

With corrplot package, we can use the cor.mtest() to compute the p-values and confidence interval for each pair of variables.

```{r}
wine.sig = cor.mtest(wine[, 1:11], conf.level = .95)

```

We can then use the p.mat argument of corrplot function as shown in the code chunk below.

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

## Reorder a corrgram

Matrix reorder is very important for mining the hiden structure and pattern in a corrgram. By default, the order of attributes of a corrgram is sorted according to the correlation matrix (i.e. “original”). The default setting can be over-write by using the order argument of corrplot(). Currently, corrplot package support four sorting methods, they are:

“AOE” is for the angular order of the eigenvectors. See Michael Friendly (2002) for details. “FPC” for the first principal component order. “hclust” for hierarchical clustering order, and “hclust.method” for the agglomeration method to be used. “hclust.method” should be one of “ward”, “single”, “complete”, “average”, “mcquitty”, “median” or “centroid”. “alphabet” for alphabetical order. “AOE”, “FPC”, “hclust”, “alphabet”. More algorithms can be found in seriation package.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")

```

## Reordering a correlation matrix using hclust

```{r}

corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
